### 2017-12-09 paper.tex
Initial commit

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Performance Factors for Large Scale Machine Learning Applications}

\author{\IEEEauthorblockN{Johannes Wünsche}
\IEEEauthorblockA{University Magdeburg\\
Email: johannes.wuensche@st.ovgu.de}%since this is University intern only probably ok otherwise this would be shady
\and
\IEEEauthorblockN{Gabriel Campero Durand}
\IEEEauthorblockA{University Magdeburg}}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
\begin{itemize}
\item Deliver short introduction for neural networks with different aspect focused examples
\item What we want to show in this paper(performance factors)
\item To what end we came
\item some other stuff that comes up during the paper and is important enough to be named at the beginning
\end{itemize}
\end{abstract}

\IEEEpeerreviewmaketitle
\section{Introduction}
% no \IEEEPARstart
\begin{itemize}
\item Deliver a short introduction coining the expressions neural network(deep learning , recurrent, deep belief ...) and explain to focus in the further paper on deep learning 
\item To what extend did we explore the performance of machine learning related to which factors and specialization of specific kind of nets
\item Show which methods were used to explore neural nets
\item Name the result again that was reached at the end
\end{itemize}
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

\subsection{Neural Nets}
A Neural Net describes an entity consisting of multiple kind of nodes, which are the inputs responsible for receiving the data belonging to the current input vector, the outputs acting as the units for the output vector in the net, and the hidden nodes responsible for transforming the data for later usage or output. These can be ordered into two groups, the open layers which contains the input and output nodes, and the hidden layers containing the hidden nodes. Although no hidden node has to be present to guarantee an acceptable result of the net for some problems and data sets.
%TODO work out this para for better condensation of content

\subsubsection{Neural Nodes}
explain the base idea of deciding unit, perceptron, sigmoid soso

\subsubsection{Step to Nets}
very short step to neural nets and why there are different types

\subsection{Implementations for different languages}%maybe not will ask gabriel
here i will shortly name all implementations relevant to the following content

\subsection{Performance of Neural Nets}
here belongs a describtion of all the observed performance measures which we will use to classify the following factors to their influence on the overall performance

\subsection{Performance Factors and their influence}
here will go the tables and their explanation of different factors found with their influence over scale of data and complexity of the net

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
yeah ... pretty much the conclusion shortly repeating the basis that is important for it and point the way they were created, should be larger part to emphasize the importance of some factors over others and the reasoning behind this classification

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}jo,
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2017-12-16 paper.tex
changes 15th December

@@
 % Do not put math or special symbols in the title.
-\title{Performance Factors for Large Scale Machine Learning Applications}
+\title{
+%Performance Factors for Large Scale Machine Learning Applications
+A Beginners Guide to Perfomance Factors of Machine Learning Learning Applications with Focus on Neural Nets
+}
 
 \author{\IEEEauthorblockN{Johannes Wünsche}
-\IEEEauthorblockA{University Magdeburg\\
-Email: johannes.wuensche@st.ovgu.de}%since this is University intern only probably ok otherwise this would be shady
-\and
-\IEEEauthorblockN{Gabriel Campero Durand}
-\IEEEauthorblockA{University Magdeburg}}
+\IEEEauthorblockA{Otto-von-Guericke University Magdeburg\\
+Email: johannes.wuensche@st.ovgu.de}
+%\and
+%\IEEEauthorblockN{Gabriel Campero Durand}
+%\IEEEauthorblockA{Otto-von-Guericke University Magdeburg}
+}
 
@@
 % no \IEEEPARstart
+BE A BIT PITCHIER!\\
+Machine Learning seems to everywhere around us.
+With the success of Deep Learning, Convolutional, Deep Belief, and other Neural Net specifications in the recent years this trend continues to be more and more influencing everyday application, may it be from music recommendation to personalized advertisement directly addressed to the consumer.
+But also in modern automation and research machine learning becomes a greater and greater field.
+We want to have a look at the most performance influencing factors while designing and implementing these networks.
+
 \begin{itemize}
@@
 
-\subsection{Neural Nets}
-A Neural Net describes an entity consisting of multiple kind of nodes, which are the inputs responsible for receiving the data belonging to the current input vector, the outputs acting as the units for the output vector in the net, and the hidden nodes responsible for transforming the data for later usage or output.
+%\subsection{Neural Nets}
+%A Neural Net describes an entity consisting of multiple kind of nodes, which are the inputs responsible for receiving the data belonging to the current input vector, the outputs acting as the units for the output vector in the net, and the hidden nodes responsible for transforming the data for later usage or output.
 These can be ordered into two groups, the open layers which contains the input and output nodes, and the hidden layers containing the hidden nodes.
 Although no hidden node has to be present to guarantee an acceptable result of the net for some problems and data sets.
-%TODO work out this para for better condensation of content
+%%TODO work out this para for better condensation of content
+%
+%\subsubsection{Neural Nodes}
+%explain the base idea of deciding unit, perceptron, sigmoid soso
+%
+%\subsubsection{Step to Nets}
+%very short step to neural nets and why there are different types
+%
+%\subsection{Implementations for different languages}%maybe not will ask gabriel
+%here i will shortly name all implementations relevant to the following content
+% 
+%\subsection{Performance of Neural Nets}
+%here belongs a describtion of all the observed performance measures which we will use to classify the following factors to their influence on the overall performance
+%
+%\subsection{Performance Factors and their influence}
+%here will go the tables and their explanation of different factors found with their influence over scale of data and complexity of the net
 
-\subsubsection{Neural Nodes}
-explain the base idea of deciding unit, perceptron, sigmoid soso
+\section{Choice of Calculation Unit}
+The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
+\subsection{CPU}
+As the first most naive option we looked on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
+\subsubsection{Single-thread}
+The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required \emph{NUMBER OF EXAMPLE CAN BE ADDED HERE}.
+\subsubsection{Multi-thread}
+A more complex approach consist of utilizing the multi-core characterisitics of modern CPUs, by splitting of training calculation with only requirements to already calculated values.
+This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm.
+%maybe a nice little table here with speed improvements of different frameworks over the amount of coures used during the calculation
+Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage[link performance paper here(believe it was benchmarking state of the art neural network frameworks)]
+\subsection{GPU}
+A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs.
+For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the naive approachs performance.
+Allmost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.
 
-\subsubsection{Step to Nets}
-very short step to neural nets and why there are different types
+Problems lie in the tightly restricted memory of graphic cards that may not be able to contain all nodes with their corresponding weights as well as training dataset.
+This leads to some communication overhead depending on the communication strategy chosen.
 
-\subsection{Implementations for different languages}%maybe not will ask gabriel
-here i will shortly name all implementations relevant to the following content
-
-\subsection{Performance of Neural Nets}
-here belongs a describtion of all the observed performance measures which we will use to classify the following factors to their influence on the overall performance
-
-\subsection{Performance Factors and their influence}
-here will go the tables and their explanation of different factors found with their influence over scale of data and complexity of the net
+\section{Communication strategy}
+Here I want clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
 
@@
 \begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}jo,
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
+\bibitem{IEEEhowto:kopka}pap,
+S. Shi, Q. Wang, P. Xu and X. Chu, \emph{Benchmarking State-of-the-Art Deep Learning Software}, HK Baptist University, Hong Kong, 17th Feb 2017
 \end{thebibliography}

### 2017-12-19 paper.tex
changes 19th December

@@
 % *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
+
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 Email: johannes.wuensche@st.ovgu.de}
-%\and
-%\IEEEauthorblockN{Gabriel Campero Durand}
-%\IEEEauthorblockA{Otto-von-Guericke University Magdeburg}
 }
@@
 % no \IEEEPARstart
-BE A BIT PITCHIER!\\
-Machine Learning seems to everywhere around us.
-With the success of Deep Learning, Convolutional, Deep Belief, and other Neural Net specifications in the recent years this trend continues to be more and more influencing everyday application, may it be from music recommendation to personalized advertisement directly addressed to the consumer.
-But also in modern automation and research machine learning becomes a greater and greater field.
-We want to have a look at the most performance influencing factors while designing and implementing these networks.
+Machine Learning seems to be everywhere around us.
+With the success of Deep , Convolutional, Deep Belief, and other Neural Net specifications in the recent years because of the increasing capabilities of modern hardware, this trend continues to be more and more influencing on everyday application, may it be from music recommendation to personalized advertisement directly addressed to the consumer.
+But also in modern automation and research machine learning becomes a greater field managing to find more efficient ways to achieve better results in many expertise.
+We want to have a look at the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scale.
 
@@
 
-\section{Choice of Calculation Unit}
+\section{Choice of Processing Unit}
 The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
@@
 This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm.
-%maybe a nice little table here with speed improvements of different frameworks over the amount of coures used during the calculation
-Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage[link performance paper here(believe it was benchmarking state of the art neural network frameworks)]
+\begin{figure}
+\centering
+\begin{tabular}{c c c}
+FW & 1 Thread & 4 Threads\\
+\end{tabular}
+\caption{Simulation results for the network.}
+\label{fig_sim}
+\end{figure}
+
+Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
 \subsection{GPU}
 A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs.
-For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the naive approachs performance.
+For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}.
 Allmost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.
@@
 \section{Communication strategy}
-Here I want clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
+Here I want to 
+clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
+TODO: Spanish Paper on SuperComputer Learning
 
@@
 pretty much the conclusion shortly repeating the basis that is important for it and point the way they were created, should be larger part to emphasize the importance of some factors over others and the reasoning behind this classification
+\cite{shi2016benchmarking}
+
+TODO: Takeaway, different ideas, innnovative
 
@@
 \section*{Acknowledgment}
-
-The authors would like to thank...
+The author would like to thank Gabriel Campero Durand for advise and help to write the paper
 
@@
 
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-\bibitem{IEEEhowto:kopka}pap,
-S. Shi, Q. Wang, P. Xu and X. Chu, \emph{Benchmarking State-of-the-Art Deep Learning Software}, HK Baptist University, Hong Kong, 17th Feb 2017
-\end{thebibliography}
+\bibliography{paper}
+\bibliographystyle{IEEEtran}
 

### 2017-12-20 paper.tex
changes to Introduction containing the structure of the paper with the division into small and large scale applications

@@
 \section{Introduction}
-% no \IEEEPARstart
 Machine Learning seems to be everywhere around us.
-With the success of Deep , Convolutional, Deep Belief, and other Neural Net specifications in the recent years because of the increasing capabilities of modern hardware, this trend continues to be more and more influencing on everyday application, may it be from music recommendation to personalized advertisement directly addressed to the consumer.
+With the recent success of Deep, Convolutional and other Neural Networks, these kind of Machine Learning tools experience a kind of renaissance and usage in applications from industrial to service oriented.
 But also in modern automation and research machine learning becomes a greater field managing to find more efficient ways to achieve better results in many expertise.
 We want to have a look at the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scale.
+At first we attend the choice of neural network and the impact of this decision.
+Afterwards we chose to have a look, on one hand, at small scale neural networks running on a single machine without large data sets and a low query frequency after the initial training, and at the other hand, large scale neural networks operating with more layers and nodes, large data sets and distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
 

### 2017-12-24 paper.tex
incorporate some changes you suggested not finished, add performance table for multithread cpu

@@
 %Performance Factors for Large Scale Machine Learning Applications
-A Beginners Guide to Perfomance Factors of Machine Learning Learning Applications with Focus on Neural Nets
+A Beginners Guide to Perfomance Factors of Machine Learning Learning Applications with Focus on Neural Nets and Deep Learning
 }
@@
 \begin{abstract}
-\begin{itemize}
-\item Deliver short introduction for neural networks with different aspect focused examples
-\item What we want to show in this paper(performance factors)
-\item To what end we came
-\item some other stuff that comes up during the paper and is important enough to be named at the beginning
-\end{itemize}
+%\begin{itemize}
+%\item Deliver short introduction for neural networks with different aspect focused examples
+%\item What we want to show in this paper(performance factors)
+%\item To what end we came
+%\item some other stuff that comes up during the paper and is important enough to be named at the beginning
+%\end{itemize}
+
+Machine Learning is probably the most present topic of computer science in the media.
+This popularity only increased over the recent years thorugh widely publicly known apllications ranging from DeepBlue(not so recent) to IBM Watson and AlphaGo.
+Because of these successes Machine Learning is sometimes viewed as omnipotent solution to every imaginable problem.
+This is most certainly not the case, Neural Nets and alike are powerful but require a large amount of maintenance
+
 \end{abstract}
@@
 Machine Learning seems to be everywhere around us.
-With the recent success of Deep, Convolutional and other Neural Networks, these kind of Machine Learning tools experience a kind of renaissance and usage in applications from industrial to service oriented.
-But also in modern automation and research machine learning becomes a greater field managing to find more efficient ways to achieve better results in many expertise.
-We want to have a look at the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scale.
+With the recent success of Deep, Convolutional,and Neural Networks, and other Machine Learning tools, especially the, in general media, well reported Learners like AlphaGo, Silver or Watson, these are experiencing a kind of renaissance their usage has spread from industrial(e.g.
+Self-driving cars rely on autonomous interpretation of camera feed and sensor output that can be optimized by using machine learning) to retail.
+But also in modern automation and research, machine learning becomes a greater field managing to find more efficient ways to achieve better results in many expertise.
+
+In this paper we want to observe and identify the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scales.
+We expect that this study will provide help for readers interested in learning about the requirements for managing data and operations of an efficient neural network.
+
+TODO: related surveys
+
  At first we attend the choice of neural network and the impact of this decision.
@@
 
-\begin{itemize}
-\item Deliver a short introduction coining the expressions neural network(deep learning , recurrent, deep belief ...) and explain to focus in the further paper on deep learning 
-\item To what extend did we explore the performance of machine learning related to which factors and specialization of specific kind of nets
-\item Show which methods were used to explore neural nets
-\item Name the result again that was reached at the end
-\end{itemize}
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
+\section{Background}
+The base of this paper is of course the model and idead of the neural network which the reader should be familiar with. Further ...
+
+\section{Choice of Neural Nets}
 
@@
 This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm.
+The effect
 \begin{figure}
 \centering
-\begin{tabular}{c c c}
-FW & 1 Thread & 4 Threads\\
+\begin{tabular}{c c c c c}
+\hline
+Framework & 1 Thread & 2 Threads &4 Threads & 8 Threads\\\hline
+Caffe & 1.324 & 0.790 & 0.578 & 15.444 \\
+Tensorflow & 7.062 & 4.789 & 2.648 & 1.938 \\
+Torch & 1.329 & 0.710 & 0.423 & na \\\hline
 \end{tabular}
-\caption{Simulation results for the network.}
-\label{fig_sim}
+\caption{Training time(in s) for a mini-batch of size 64 of a Fully Connected Network(FCN) trained with synthetic data on a i7-3820 with 4-physical cores using 3 different commonly used Deep Learning frameworks\cite{shi2016benchmarking}}
+\label{fig_ttfcn}
 \end{figure}
@@
 
-\section{Communication strategy}
+\section{Scaling challenges}
 Here I want to 
 clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
-TODO: Spanish Paper on SuperComputer Learning
+TODO: Spanish thesis on SuperComputer Learning
 

### 2018-01-04 paper.tex
small update 4.1

@@
 \section{Background}
-The base of this paper is of course the model and idead of the neural network which the reader should be familiar with. Further ...
+The reader is advised to be familiar with the base idea of a neural network and some common types of networks like the deep, convolutional, feed forward and recurrent networks.
+
+The most common neural network frameworks which we, because of their wide popularity, will mostly view are Tensorflow\cite{abadi2016tensorflow}, Caffe\cite{jia2014caffe}, Torch\cite{collobert2002torch}, CNTK, deeplearning4j, Keras, MXNET.
+These frameworks have been benchmarked in numerous settings and applications and can help us show performance influencing factors.
 
 \section{Choice of Neural Nets}
-
-%\subsection{Neural Nets}
-%A Neural Net describes an entity consisting of multiple kind of nodes, which are the inputs responsible for receiving the data belonging to the current input vector, the outputs acting as the units for the output vector in the net, and the hidden nodes responsible for transforming the data for later usage or output.
-These can be ordered into two groups, the open layers which contains the input and output nodes, and the hidden layers containing the hidden nodes.
-Although no hidden node has to be present to guarantee an acceptable result of the net for some problems and data sets.
-%%TODO work out this para for better condensation of content
-%
-%\subsubsection{Neural Nodes}
-%explain the base idea of deciding unit, perceptron, sigmoid soso
-%
-%\subsubsection{Step to Nets}
-%very short step to neural nets and why there are different types
-%
-%\subsection{Implementations for different languages}%maybe not will ask gabriel
-%here i will shortly name all implementations relevant to the following content
-% 
-%\subsection{Performance of Neural Nets}
-%here belongs a describtion of all the observed performance measures which we will use to classify the following factors to their influence on the overall performance
-%
-%\subsection{Performance Factors and their influence}
-%here will go the tables and their explanation of different factors found with their influence over scale of data and complexity of the net
+what net when to use, and starting values of neural nets like amount of layers and starting weights
 
@@
 \subsubsection{Single-thread}
-The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required \emph{NUMBER OF EXAMPLE CAN BE ADDED HERE}.
+The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required .
+But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Fig.
+1).
 \subsubsection{Multi-thread}
@@
 This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm.
-The effect
+The effect  continues to grow by using more threads but reaches it limits when using more threads than physical cores available.
 \begin{figure} 
@@
 
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.2\textwidth]{a.png}
+%
+%\caption{Time per mini-batch in s compared between different amount of GPUs used\cite{shi2016benchmarking}}
+%\label{fig_m_gpu}
+%\end{figure}
+
+\subsection{GPU-Cluster}
+The next logical step is to use multiple GPUs further reducing training time but also creating new challenges, because of the decentralized nature of the Cluster, like communication overhead.
+This is mostly used in large scale applications and is most of the time not needed in small scale.
+
+A main advantage of the Cluster is of course the increased computation power available during the training phase, reducing training time by 35\% when doubling the number of GPUs in CNTK and MXNET, 28 \% in Caffe and 10 \% in  Torch and Tensorflow\cite{shi2016benchmarking}.
+
+Another advantage is the greater availability of memory somewhat easing the restrictions of the memory restricted GPUs structure.
+While this is an advantage it is not efficient if used as the single goal of multiple GPU usage because of the high price connected to acquiring them.
+
+Fig.
+2 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs used.
+Although costly this leads to a constant improvement.
+Higher Numbers of GPUs can be again slower or offer no significant improvement if used the same as smaller amounts because of an increasing communication overhead and bottleneck in the CPU, which is responsible for managing seperate GPUs.
+
+\subsubsection{Multiple GPU-Clusters}
+Taking a step further it is possible to connect multiple Clusters of GPUs to compute a single task.
+In doing this we again increase raw computation power, but also increase the total amount of communication required.
+
 \section{Scaling challenges}
-Here I want to 
-clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
+Here I want to clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
 TODO: Spanish thesis on SuperComputer Learning
@@
 
+\bibliographystyle{IEEEtran}
 \bibliography{paper}
-\bibliographystyle{IEEEtran}
 

### 2018-01-06 paper.tex
add Background and Related Work, citing not final, enhanced bibtex

@@
 
-\section{Background}
-The reader is advised to be familiar with the base idea of a neural network and some common types of networks like the deep, convolutional, feed forward and recurrent networks.
+\section{Background and Related Work}
+
+%%%%BEGIN SMALL SCALE NEURAL NET DESCRIPTION
+
+As a small scale neural network applications we define a network with few neural nodes inside the network and because of this rather limited size most of the time run on a single machine and or single processing unit(worker) which can complete the training set in a reasonable time.
+
+%%%%END SMALL SCALE NEURAL NET DESCRIPTION
+
+%%%%BEGIN LARGE SCALE NEURAL NET DESCRIPTION
+
+On the other hand as a large scale neural network applications we define a network with a large number of nodes which can because of the complexity of the resulting connections take a large amount of time to complete training if computed on one worker and therefore will take use of a distributed system consisting of multiple workers and one or more servers to synchronize the workers and distribute the required data.
+
+%%%%END  LARGE SCALE NEURAL NET DESCRIPTION
+
+%%%%BEGIN DESCRIPTION NETWORKS USED LATER
+
+Most commonly used Neural Networks of the more classic single or non hidden layer kind are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a simple layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
+
+Feed Forward Networks(FFN) simply connects input, hidden and output layers with an all to all connection between the layers with the number of outputs being equal or smaller to the number of hidden nodes.
+
+Auto Encoders(AE) are of an similar build but the hidden layer compresses the input information into fewer nodes while these are later again decompressed to output nodes that match the previously given input nodes.
+
+Restricted Boltzmann Machines(RBM) only have an input and an output layer with an all to all connection between these.
+This RBM has in comparison to the normal Boltzmann Machine a better training behaviour which leads to a faster trained network.
+
+A more advanced architecture is to increase the number of hidden layers to enlarge the application possibility of these networks for more complex models these are then called Deep Networks.
+The most commonly used ones are Variational Auto Encoders , Deep Belief Networks, Generative Adversial Networks, Recurrent Neural Networks ,e.g.
+Long short-term memory networks, Deep Convolutional Neural Networks, Deconvolutional Networks.
+
+Variational Auto Encoders(VAE) consists of a encoder, sampler and decoder and are able to to generate output based on specification given to the VAE.
+
+Deep Belief Networks(DBN) are a combination of RBM and FFN in which the RBM are used to pretrain the network with the initial training 
+data and the FFN to adjust the learned values slightly to improve the result.
+
+Generative Adversial Networks(GAN) like the VAE creates random output based on the model found in the given training data.
+VAE are a specialization of GAN because of the restrictions that can be set for the output which is not possible in a GAN.
+
+Deep Convolutional Networks(DCN) are based on Convolutional Nodes compressing the incoming information into the needed base model and then calculate the required output.\cite{PattersonGibson17} They are used to abstract input information into more general models.
+
+Deconvolutional Networks(DN)\cite{zeiler2014visualizing} are less surprisingly the opposite of DCN in the idea that input information is first spread to a larger amount of nodes and then processed by the following Nodes to be used by the output.
+
+Long short-term memory(LSTM)\cite{hochreiter1997long} are a variant of Recurrent Neural Networks including additionally memory cells and gates that control the content of the cells.
+This model has been able to improve learning process of RNN and results.
+
+%%%%END DESCRIPTION NETWORKS USED LATER
+
+%%%%BEGIN TRAINING ALGORITHMS
+%%THE POSITION OF THIS CAN BE CHANGED
+The reason why the training of neural networks is so expensive can be traced back to the algorithm used to train all commonly used networks, which is backpropagation.
+Backpropagation is based on the idea that we calculate the deviation of the output values from the actual result and then use this backwards thorugh the layers of the net to calculate a correction,based on the error of the node and the learning rate, for each connection weight until the input layer is reached\cite{riedmiller1993direct}.
+Some Networks like DBN use a slightly
+diverged "Gentle" Backpropagation, which uses a lowered learning rate after the initial values have been found from the training data with the help of the pretrain RBM \cite{PattersonGibson17}.
+%%%%END TRAINING ALGORITHMS
+
+%%%%BEGIN INFORMATION DATA FLOW
+
+Another performance influencing factor besides the actual neural net and the used learning algorithm is the amount of data required for the net to be trained and the size of the input of each training example.While this is mostly important for large scale applications it can also be hindering for smaller ones.
+For example if we use a Convolutional Neural Network to classify the content of a greyscale picture we get for small $512*512$ sized images $262144$ inputs that each need to be transferred to the processing unit, CPU or GPU, for each example.
+
+This large amount of data can for one take some time to be transferred, if for example this data package is required by an external processing unit like a GPU other units can not  receive new data and therefore lose performance due to them being idle.
+Also if too much data is needed in relation to the actual output the data storage can achieve enduring performance loss will occur.
+
+Further can the required training data exceed direct accessible memory from the processing unit, enhancing the problem of delivering data to all processing units.
+
+Additionally the synchronization can be somewhat difficult and extensive for large neural networks on multiple processing units because of the interchange of new weights calculated between them after a certain number of iterations on their training cases.
+
+%%%%END INFORMATION DATA FLOW
+
+%%%%BEGIN INFORMATION ACTUAL USED IMPLEMENTATIONS
 
@@
 These frameworks have been benchmarked in numerous settings and applications and can help us show performance influencing factors.
+
+%TODO some more detailed stuff about it
+%%%%END INFORMATION ACUTAL USED IMPLEMENTATIONS
+
+%%%%BEGIN RELATED WORK
+
+The topic of performance of neural networks and influencing factors has been touched by many benchmarking survey like \cite{shi2016benchmarking} and \cite{qi2016paleo}.	 But we hope to offer a collected view on the most influencing factors in a more abstract style.
+
+%%%%END RELATED WORK
 

### 2018-01-06 paper.tex
add Choice of Neural Network section and further citing in background(still not final though)

@@
 
+%%%%IMAGE PACKAGE
+\usepackage{graphicx}
+
 % *** FLOAT PACKAGES ***
@@
 
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
+\begin{document}
 
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
 \title{
@@
 \begin{abstract}
-%\begin{itemize}
-%\item Deliver short introduction for neural networks with different aspect focused examples
-%\item What we want to show in this paper(performance factors)
-%\item To what end we came
-%\item some other stuff that comes up during the paper and is important enough to be named at the beginning
-%\end{itemize}
 
@@
 The most commonly used ones are Variational Auto Encoders , Deep Belief Networks, Generative Adversial Networks, Recurrent Neural Networks ,e.g.
-Long short-term memory networks, Deep Convolutional Neural Networks, Deconvolutional Networks.
+Long short-term memory networks, Deep Convolutional Neural Networks, Deconvolutional Networks, and Recursive Neural Networks.
 
@@
 
-The most common neural network frameworks which we, because of their wide popularity, will mostly view are Tensorflow\cite{abadi2016tensorflow}, Caffe\cite{jia2014caffe}, Torch\cite{collobert2002torch}, CNTK, deeplearning4j, Keras, MXNET.
+The most common neural network frameworks which we, because of their wide popularity, will mostly view are Tensorflow\cite{abadi2016tensorflow}, Caffe\cite{jia2014caffe}, Torch\cite{collobert2002torch}, CNTK\cite{gitcntk}, deeplearning4j\cite{websitedl4j}, Keras\cite{websitekeras}, MXNET\cite{websitemxnet}.
 These frameworks have been benchmarked in numerous settings and applications and can help us show performance influencing factors.
@@
 %%%%END INFORMATION ACUTAL USED IMPLEMENTATIONS
+
+%%%%BEGIN PROCESSING UNIT
+
+TODO
+
+%%%%END PROCESSIONG UNIT
 
@@
 
-\section{Choice of Neural Nets}	
-what net when to use, and starting values of neural nets like amount of layers and starting weights
+\section{Choice of Neural Network}
+
+%%%%BEGIN IMPORTANCE
+The first important choice is to what neural network is going to be used.
+This decision is heaviliy reliant on the application that is to be implemented.
+This decision influences all further actions and has to be carefully made, to assist this decision we will offer some generalization when to use simple neural networks or deep neural network and what specializations of certain neural networks are giving them advantage while doing specific tasks.
+%%%%END IMPORTANCE
+
+%%%%BEGIN ADVANTAGES DISADVANTAGES SIMPLE AND DEEP
+\subsection{What kind of networks to use}
+\subsubsection{Simple Networks}
+Simple task can be done by a network which is not deep like a FFN or a RBM.
+They are faster to train and even larger training sets can be completed in a reasonable time without multithreading or usage of GPUs.
+But there small size can lead that accuracy decreses when the model becomes more difficult than anticipated.
+Small scale applications often incur that a simple neural network is used because of the smaller training set and shorter computation time.
+
+\subsubsection{Deep Networks}
+More complex tasks more often require a deep neural network that can learn a more complex model than simple networks and offer a greater accuracy while doing this.
+The major downside of these networks that they require a larger amount of time to train and can be oversized for the actual to be represented model, and therefore adding unnecessary computation time.
+Large scale applications mostly contain deep networks because of their larger training sets and more complex models underlying the actual task.
+
+We can not offer a clear answer when to use which kind of network, every problem must be individually analyzed and based on the information, and complexity, as well as the acceptable error rate of the final network the decision has to be made.
+For cases that are unclear a simple network can first be constructed,because of their short training time, and if the result is unsatisfying a more complex network can be considered.
+%%%%END ADVANTAGES DISADVANTAGES SIMPLE AND DEEP
+
+%%%%INTITAL VALUES AND AMOUNT OF NODES
+
+An also important choice is the setting of the initial values of connection weights, here it is advisable to start at low values like $0.1$ to prevent straying too far from the optimal solution, since algorithm like backpropagation mostly find a local optimum initial values too far of from the global optimum will deliver worse results.
+
+%%%%END INITIAL VALUES AND AMOUNT OF NODES
+
+%%%%DETAILS SIMPLE AKA SMALL
+\subsection{Simple network Nodes}
+If a simple network shall be used the next choice will be the amount of nodes in the hidden layer.
+This is of course dependent one foremost the number of inputs and outputs, the number of hidden nodes should deviate to strong from them.
+How strong this deviation really is is dependent on the chosen neural network, for example a feed forward network profits when the number is equal or larger than the number of inputs, while an auto encoder probably has fewer hidden nodes.
+For a simple network this decision can be if too far from the optimum less fatal than for a deep network, while a simple network may lose some accuracy or gain additional training time a deep network can experience these effects at a greater scale.
+
+%%%%END DETAILS SIMPLE AKA SMALL
+
+%%%%DETAILS DEEP AKA LARGE AND SMALL
+\subsection{Deep networks Nodes}
+Like in simple networks the amount of nodes to be used is an important value that determines the final ability of the network to fulfill its task in the required accuracy and speed.
+The acutal number of nodes chosen at the end is of course a bit more complex in deep networks for example networks like DCN require a special node configuration based on their basic idea(step by step decreasing of nodes per layer until desired compression reached) , or DN(opposite of DCN).
+So how the node configuration looks like is strongly dependent on the net chosen.
+
+%%%%END DETAILS DEEP AKA LARGE AND SMALL
+
+%%%%SPECIFIC NETS
+
+\subsection{Specific deep networks}
+
+%%%%END SPECIFIC NETS
 
@@
 \end{figure}
-
 Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.2\textwidth]{a.png}
-%
-%\caption{Time per mini-batch in s compared between different amount of GPUs used\cite{shi2016benchmarking}}
-%\label{fig_m_gpu}
-%\end{figure}
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{a.png}
+
+\caption{Time per mini-batch in s compared between different amount of GPUs used\cite{shi2016benchmarking}}
+\label{fig_m_gpu}
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{gpu_cluster_perf_win.png}
+\caption{Time in h of different asynchronous configurations until reaching destined error rate(loss)\cite{sastre2017scalability}}
+\label{fig_cl_gpus}
+\end{figure}
+
 \subsubsection{Multiple GPU-Clusters}
@@
 In doing this we again increase raw computation power, but also increase the total amount of communication required.
+Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters (Fig.
+3)
 
@@
 TODO: Spanish thesis on SuperComputer Learning
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
 

### 2018-01-07 paper.tex
update sunday 7th january

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN BACKGROUND
+
 \section{Background and Related Work}
@@
 
-Long short-term memory(LSTM)\cite{hochreiter1997long} are a variant of Recurrent Neural Networks including additionally memory cells and gates that control the content of the cells.
+Long short-term memory(LSTM)\cite{hochreiter1997long} are a variant of Recurrent Neural Networks(RNN) including additionally memory cells and gates that control the content of the cells.
 This model has been able to improve learning process of RNN and results.
@@
 
-TODO
+The next major performance influencing factor will be the choice of the actual processing unit.
+This decision is the most computation time influencing and can also influence the quality of the result.
+We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but posses a smaller virtualization limit than GPUs because of their differentiating architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters,being simple a seperate unit consisting of multiple GPUs, like they are used in supercomputers.
 
@@
 %%%%END RELATED WORK
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CHOICE OF NETWORK
 
@@
 For cases that are unclear a simple network can first be constructed,because of their short training time, and if the result is unsatisfying a more complex network can be considered.
+Also it is advisable to use a deep network if the data is either of image or audio nature and/or needs to include changes over time of an input signal\cite{PattersonGibson17}.
 %%%%END ADVANTAGES DISADVANTAGES SIMPLE AND DEEP
 
-%%%%INTITAL VALUES AND AMOUNT OF NODES
+%%%%INTITAL VALUES
 
@@
 
-%%%%END INITIAL VALUES AND AMOUNT OF NODES
+%%%%END INITIAL VALUES
 
@@
 \subsection{Specific deep networks}
+Based on the content of the task a few networks can be more fitting than others to fulfill it.
+To deliver a shot overview of it we use a classification by Patterson\cite{PattersonGibson17} for the basis of our advise.
 
+If the task includes the generation of data to be output a GAN, VAE or RNN/LSTM.
+Which of these to choose is strongly dependent of the specific parameters of the task.
+For example if we want to generate random new data following the model of our training data the closest to this would be a GAN because it is by design designated to fulfill such tasks.
+But if new data which is not completely random is to be generated a VAE can be considered which offers by design this possibility.
+In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} or text generation with LSTM\cite{websitetext}.%this citation is broken need to fix later
+
+If the task includes classification or interpretation of visual data neural networks like CNN or DBN can be used.
+For once CNN are a straightforward choice since they are able to compress data and therefore abstract smaller data like the content of an image from the original image.
+But DBN offer also an possibility to complete the task with the major adavantage of being able to first roughly guess values from the training data and then train unsupervised in the fine-tuning phase as shown by Zhong 2016\cite{zhong2016diversified}.
+
+If the task includes the classification or intepretation of incoming data over time the ability of RNN and LSTMs to data that is of sequential nature comes in handy.
+For example voice recognition is possible to implement with the help of LSTMs as shown by Soltau 2016\cite{soltau2016neural}.
 %%%%END SPECIFIC NETS
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CHOICE OF PROCESSING UNIT
 
@@
 \subsection{CPU}
-As the first most naive option we looked on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
+As the first most naive option we have a look on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
 \subsubsection{Single-thread}
@@
 In doing this we again increase raw computation power, but also increase the total amount of communication required.
-Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters (Fig.
-3)
+Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in [Fig.
+3]
 
-\section{Scaling challenges}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CHALLENGES
+
+\section{Large scale challenges vs Small scale challenges}
 Here I want to clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
 TODO: Spanish thesis on SuperComputer Learning
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CONCLUSION
 

### 2018-01-08 paper.tex
fix one broken citation and added one to examples

@@
 But if new data which is not completely random is to be generated a VAE can be considered which offers by design this possibility.
-In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} or text generation with LSTM\cite{websitetext}.%this citation is broken need to fix later
+In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017}(implementing theory by Zhu 2016\cite{zhu2016generative}) or text generation with LSTM\cite{gittesttensorflow}.%this citation is broken need to fix later
 

### 2018-01-08 paper.tex
fix one citation

--- 2018-01-07 paper.tex
@@
 But if new data which is not completely random is to be generated a VAE can be considered which offers by design this possibility.
-In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} or text generation with LSTM\cite{websitetext}.%this citation is broken need to fix later
+In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} (corresponding paper \cite{zhu2016generative}) or text generation with LSTM\cite{gittesttensorflow}.
 

### 2018-01-08 paper.tex
addition 8th January

--- 2018-01-07 paper.tex
@@
 But if new data which is not completely random is to be generated a VAE can be considered which offers by design this possibility.
-In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} or text generation with LSTM\cite{websitetext}.%this citation is broken need to fix later
+In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} (corresponding paper \cite{zhu2016generative}) or text generation with LSTM\cite{gittesttensorflow} (paper describing base approach\cite{sutskever2011generating}).
 
@@
 The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required .
-But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Fig.
-1).
+But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Table I).
 \subsubsection{Multi-thread}
@@
 The effect  continues to grow by using more threads but reaches it limits when using more threads than physical cores available.
-\begin{figure} 
+\begin{table} 
 \centering
@@
 \label{fig_ttfcn}
-\end{figure}
+\end{table}
 Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
@@
 
+\subsection{Multi-GPU}
+To further improve performance during training multiple GPUs can be used to train the neural net.
+This shortens training time but also requires new actions like synchronizations of GPUs and splitting of data and/or neural network.
+
+A main advantage of Multi-GPU is of course the increased computation power available during the training phase, reducing training time by 35\% when doubling the number of GPUs in CNTK and MXNET, 28 \% in Caffe and 10 \% in  Torch and Tensorflow\cite{shi2016benchmarking}.
+
+Another advantage is the greater availability of memory somewhat easing the restrictions of the memory restricted GPUs structure.
+While this is an advantage it is not efficient if used as the single goal of multiple GPU usage because of the high price connected to acquiring them.
+
+Fig.
+1 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs used.
+Although costly this leads to a constant improvement.
+Higher Numbers of GPUs can be again slower or offer no significant improvement if used the same as smaller amounts because of an increasing communication overhead and bottleneck in the CPU, which is responsible for managing seperate GPUs.
+
+\subsubsection{Parallelism between GPUs}
+There are two kind of parallelism approaches applicable to the problem, first there is data parallelism which splits the training data into multiple sets, mostly used when the training set is splitable into multiple subsets.
+This can be done by dividing the data by case or by feature.
+Second model parallelism in which the net is divided into separately runnable parts which later can be joined together again, this method is fitting for single system distribution in which the GPUs share one PCIe bus, but lead to the result being more vulnerable to calculation failures \cite{sastre2017scalability}.
+
 \begin{figure}
@@
 
-\subsection{GPU-Cluster}
-The next logical step is to use multiple GPUs further reducing training time but also creating new challenges, because of the decentralized nature of the Cluster, like communication overhead.
+\subsection{GPU-Clusters}
+
+The next logical step is to use multiple Multi-GPU systems further reducing training time but also creating new challenges, because of the decentralized nature of the Cluster, like communication overhead.
 This is mostly used in large scale applications and is most of the time not needed in small scale.
-
-A main advantage of the Cluster is of course the increased computation power available during the training phase, reducing training time by 35\% when doubling the number of GPUs in CNTK and MXNET, 28 \% in Caffe and 10 \% in  Torch and Tensorflow\cite{shi2016benchmarking}.
-
-Another advantage is the greater availability of memory somewhat easing the restrictions of the memory restricted GPUs structure.
-While this is an advantage it is not efficient if used as the single goal of multiple GPU usage because of the high price connected to acquiring them.
-
-Fig.
-2 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs used.
-Although costly this leads to a constant improvement.
-Higher Numbers of GPUs can be again slower or offer no significant improvement if used the same as smaller amounts because of an increasing communication overhead and bottleneck in the CPU, which is responsible for managing seperate GPUs.
 
@@
 \includegraphics[width=0.4\textwidth]{gpu_cluster_perf_win.png}
-\caption{Time in h of different asynchronous configurations until reaching destined error rate(loss)\cite{sastre2017scalability}}
+\caption{Time (in h) of different asynchronous configurations until reaching destined error rate(loss)\cite{sastre2017scalability}}
 \label{fig_cl_gpus}
@@
 In doing this we again increase raw computation power, but also increase the total amount of communication required.
-Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in [Fig.
-3]
+Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in (Fig.
+2)
 
@@
 
-\section{Large scale challenges vs Small scale challenges}
-Here I want to clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
-TODO: Spanish thesis on SuperComputer Learning
+\section{Large scale challenges and Small scale challenges}
+Dependent on the choice of scale each one has its own challenges to be considered during design and implementation of the neural network.
+
+\subsection{Small scale challenges}
+First and foremost it is important to notice that the biggest restrictions of small scale applications is the limited computation power available during the initial training process, later during usage of the net this is less impeding because either the net is completely pretrained and not changed later or the only change is some optimization when the network receives a query.
+
+This leads to the focus lying on the design to find the best fit of complexity to fit the desired model. Otherwise longer training time or less accuracy of the resulting network will conclude.
+
+Further small scale applications are relatively free of performance related complications, because of their relative simplicity and fast modern computer hardware enabling operations which are not optimal to be resolved in a non less desirable time
+
+\subsection{Large scale challenges}
+Opposing to small scale applications large scale ones to not have the actual processing power as their highest pending restriction but the communication and data transfer between all acting processing units.
+This problem occurs because a larger data flow has to be managed to more processing units compared to small scale applications.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{example_configuration.png}
+\caption{One example configuration consisting of a CPU and 4 connected GPU units for training the neural network.}
+\end{figure}
+
+The data flow is a problem on multiple fronts.
+For one the transfer of the training data from main memory to GPU memory can be the first bottleneck restricting faster processing speeds from being realized.
+For example if the GPU only receives data if no calculation is currently active this slows down the training process unnecessarily.
+To increase the speed of the calculation data can be fetched during the processing of already present data, this is called data prefetching as described and implemented by Yang 2010 \cite{yang2010gpgpu}.
+This solution can solve the problem partially but reaches it limits if used with multiple GPUs for example configuration (Fig.
+3), which leads that the CPU will become a bottleneck if the training data of each individual GPU is too large to fit in the GPU memory, since the CPU can only manage transfer to the GPUs one at a time.
+For example a GPU equipped with PCI-Express 4.0 with a x16 Throughput can at best receive 31.5 GB/s, if we assume that all training data can be fit into main memory, that way if we need to refresh the complete memory of a modern GPU like NVIDIA GTX 1080, we need 0.253968 s, so to refresh all GPUs memory without including any overhead produced by switching operations between the GPUs we have an absolute transfer time of 1.015873 s, and for 3 units 0.7619 s, so if our GPU completes the desired calculation faster, for example 0.5 s, we have an 0.2619 s period in which our GPU idles every iteration, assumed that we can transfer all data during the calculation time.
+Therefore we have an effective calculation computation time each phase of 65.62 \%.
+This problem only enlargens if the main memory is to small to fit the complete training data since connections to solid state disk are too slow and hard disk do physically do not reach the required throughput.
+
+The other kind of data flow is required to synchronize the calculated neural network between the GPUs. This can be done multiple ways first
+
+The next problem coming up with multiple processing units is the data flow between these, synchronizing the seperately calculated nets to reach a united model again, also have to be managed by the CPU, hence further reducing the effective calculation time of the GPUs.
+To ease the effect of this GPU to GPU communication technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system.
 

### 2018-01-09 paper.tex
add conclusion, reworked abstract, reworked challenges, reworked processing unit

@@
 
-Machine Learning is probably the most present topic of computer science in the media.
-This popularity only increased over the recent years thorugh widely publicly known apllications ranging from DeepBlue(not so recent) to IBM Watson and AlphaGo.
+Machine Learning is probably the most present topic of computer science in general media.
+This popularity only increased over the recent years thorugh widely publicly known applications ranging from DeepBlue(not so recent) to IBM Watson and AlphaGo.
 Because of these successes Machine Learning is sometimes viewed as omnipotent solution to every imaginable problem.
-This is most certainly not the case, Neural Nets and alike are powerful but require a large amount of maintenance
+This is most certainly not the case, Neural Nets and alike are powerful but require a large amount of design and implementation work.
+We want to give a short introduction to performance influencing factorts of neural networks during training and the resulting network.
 
@@
 
-In this paper we want to observe and identify the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scales.
-We expect that this study will provide help for readers interested in learning about the requirements for managing data and operations of an efficient neural network.
+In this paper we want to observe and identify the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scale applications.
+We expect that this study will provide help for readers interested in learning about the the initial design of neural networks, requirements for managing data, and operations of an efficient neural network.
 
@@
 Afterwards we chose to have a look, on one hand, at small scale neural networks running on a single machine without large data sets and a low query frequency after the initial training, and at the other hand, large scale neural networks operating with more layers and nodes, large data sets and distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
+We divided our survey into first the choice of neural network afterwards choice of processing units, including basic data splitting strategies for multi processing units systems, followed by challenges occuring when using large and small scale applications.
 
@@
 
-\section{Choice of Processing Unit}
+\section{Choice of Processing Units}
 The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
@@
 
+\subsubsection{Synchronization between GPUs}
+For synchronizing Multi-GPU setups like in Figure 3 there are theoretically two choices.
+One being the synchronous method in which after a value is calculated they are shared with every other GPU.
+This method is applicable to all Multi-GPU configuration and delivers an acceptable result\cite{wang2016deep}.
+The update of values for all GPUs occurs only after the updated values have been send by each of them.
+Until a large amount of over 40 GPUs are used this mehtod can be used without limitation\cite{sastre2017scalability}.
+There is also the possibility to run the GPUs asynchronously, by letting them use different versions of the values and only updating them periodically by sending an update request to the CPU.
+This can lead to a slight improvement of the training speed but is not guaranteed depending on the training model\cite{wang2016deep}.
+
 \begin{figure}
@@
 
+\subsubsection{Parallelism of multiple GPU-Clusters}
+Like the Multi-GPU setup this can be done by creating either data or model parallelism.
+Combination are also possible when using GPU-Clusters, for instance in a setup consisting of multiple Clusters and one single communication server a model parallelism can be used to distribute the model on the different clusters, but they internally use data parallelism to split work\cite{wang2016deep}.
+
+\subsubsection{Synchronization of multiple GPU-Clusters}
+Similar to parallelism the synchronization of GPU-Clusters can be undertaken either with the synchronous or asynchronous like in Multi-GPU configurations.
+Again these tactics can be combined to create a mixed-asynchronous strategy, using an asynchronus training on inter-clusters level and cluster internal synchronous training\cite{sastre2017scalability}.
+This creates the best performance on large configurations with many clusters\cite{wang2016deep}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CHALLENGES
@@
 
-Further small scale applications are relatively free of performance related complications, because of their relative simplicity and fast modern computer hardware enabling operations which are not optimal to be resolved in a non less desirable time
+Further small scale applications are relatively free of performance related complications, because of their relative simplicity and fast modern computer hardware enabling operations which are not optimal to be resolved in a non less desirable time.
 
@@
 
+%%Data flow of training data
 The data flow is a problem on multiple fronts.
@@
 
-The other kind of data flow is required to synchronize the calculated neural network between the GPUs. This can be done multiple ways first
+%%data flow synchronization
+Additional data flow is created when synchronizing GPUs with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations.
 
-The next problem coming up with multiple processing units is the data flow between these, synchronizing the seperately calculated nets to reach a united model again, also have to be managed by the CPU, hence further reducing the effective calculation time of the GPUs.
-To ease the effect of this GPU to GPU communication technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system.
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{cluster_setup.png}
+\caption{One example configuration consisting of a communication server and 4 connected GPU-Clusters each containing 4 GPUs.}
+\end{figure}
+%synchronous
+Synchronous training, while being recommend for setups that have 40 GPUs or less, can create idling time if one unit is slower than others, since all updates have to be send to the server before the next iteration can be undertaken this slows down all other units.
+This problem becomes more probable the more units are used.
+Also more GPUs lead to more communication if synchronization is required after very step, so they increase the amount of data managed by the CPU or communication server.
+This applies to configuration like Figure 3 and Figure 4.
+
+%asynchronous
+If an asynchronous training approach is used problems may occur influencing the result, if updates are to sporadic and different versions of values drift too far apart.
+To avoid this problem GPU-Cluster configurations do not rely on complete asynchronous training startegies, but on the already described mixed-asynchronous minimizing this problem while still reducing possible communication overhead.
+
+To ease the effect of GPU to GPU communication,by relieveing the CPU of managing the , technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system.
 
@@
 \section{Conclusion}
-yeah ...
-pretty much the conclusion shortly repeating the basis that is important for it and point the way they were created, should be larger part to emphasize the importance of some factors over others and the reasoning behind this classification
-\cite{shi2016benchmarking}
+During our survey we observed that the most influencing factors are the basic choice of neural network to be used, this includes its complexity as well as the actual architecture the network is build of.
+Further the processing unit that is chosen has a strong influence on the quality and speed the final network that lies within acceptable error limits will be reached.
+Based on the choice of this other factors come into consideration influencing the process like communication between processing units and parallelism of data between them.
+They are not quite as influential as the first two but can lead to a strong improvement or worsening.
+
+%Takeaway
+Especially data transfer time and alike carries a potential to lengthen the training time of an significant amount.
+To ease this technology like GPUDirect\cite{nvidiagpudirect2017} and data prefetching\cite{yang2010gpgpu} can be used.
+It may be possible to decrease the effect of the data flow further by introducing a memory transfer unit which can indepently organize data and can handle parallel transfers to and from multiple units.
+This could decrease the calculation time to a bare minimum but would requires additional hardware and maybe modification of the used GPU and CPU.
+Although this would break security of the system on broder range because of the unauthorized memory access and changes, it could offer a steep improvement for large calculations on distributed systems.
 

### 2018-01-09 paper.tex
modifications made

@@
 For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}.
-Allmost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.
+Almost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.
 
@@
 Additional data flow is created when synchronizing GPUs with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations.
+But both create an communication overhead because during transmission of values the GPU cannot operate on the neural network training.
 
@@
 
-To ease the effect of GPU to GPU communication,by relieveing the CPU of managing the , technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system.
+To ease the effect of GPU to GPU communication,by relieveing the CPU of managing the , technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead
 

### 2018-01-09 paper.tex
fix heading

@@
 %Performance Factors for Large Scale Machine Learning Applications
-A Beginners Guide to Perfomance Factors of Machine Learning Learning Applications with Focus on Neural Nets and Deep Learning
+A Beginners Guide to Perfomance Factors of Machine Learning Applications with Focus on Neural Nets and Deep Learning
 }
@@
 \section{Conclusion}
+%Summary
 During our survey we observed that the most influencing factors are the basic choice of neural network to be used, this includes its complexity as well as the actual architecture the network is build of.
@@
 To ease this technology like GPUDirect\cite{nvidiagpudirect2017} and data prefetching\cite{yang2010gpgpu} can be used.
-It may be possible to decrease the effect of the data flow further by introducing a memory transfer unit which can indepently organize data and can handle parallel transfers to and from multiple units.
+It may be possible to decrease the effect of the data flow further by introducing a direct memory transfer unit which can indepently organize data and can handle parallel transfers to and from multiple units.
 This could decrease the calculation time to a bare minimum but would requires additional hardware and maybe modification of the used GPU and CPU.

### 2018-01-09 paper.tex
some changes conclusion and challenges

@@
 \IEEEauthorblockA{Otto-von-Guericke University Magdeburg\\
-Email: johannes.wuensche@st.ovgu.de}
+johannes.wuensche@st.ovgu.de}
 }
@@
 Machine Learning seems to be everywhere around us.
-With the recent success of Deep, Convolutional,and Neural Networks, and other Machine Learning tools, especially the, in general media, well reported Learners like AlphaGo, Silver or Watson, these are experiencing a kind of renaissance their usage has spread from industrial(e.g.
+With the recent success of Neural Networks, and other Machine Learning tools, especially the, in general media, well reported learners like AlphaGo, Silver or IBM Watson, these are experiencing a kind of renaissance their usage has spread from industrial(e.g.
 Self-driving cars rely on autonomous interpretation of camera feed and sensor output that can be optimized by using machine learning) to retail.
@@
 
-As a small scale neural network applications we define a network with few neural nodes inside the network and because of this rather limited size most of the time run on a single machine and or single processing unit(worker) which can complete the training set in a reasonable time.
+As a small scale neural network applications we define a network with few neural nodes inside the network and because of this rather limited size it is most of the time run on a single machine and or single processing unit which can complete the training process of the net in a reasonable time.
 
@@
 
-On the other hand as a large scale neural network applications we define a network with a large number of nodes which can because of the complexity of the resulting connections take a large amount of time to complete training if computed on one worker and therefore will take use of a distributed system consisting of multiple workers and one or more servers to synchronize the workers and distribute the required data.
+On the other hand as a large scale neural network applications we define a network with a large number of nodes which can because of the complexity of the resulting connections take a large amount of time to complete training if computed on one processing unit and therefore will take use of a distributed system consisting of multiple units in a single system or more servers to synchronize multiple systems of processing units and distribute the required data.
 
@@
 A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs.
-For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}.
-Almost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.
+For example with the help of  NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}.
+Almost all recent neural net learning frameworks support GPUs because of its excellent performance rating for neural networks.
 
@@
 
-To ease the effect of GPU to GPU communication,by relieveing the CPU of managing the , technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead
+To ease the effect of GPU to GPU communication,by relieveing the CPU of managing them, technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead.
+Also shared memory between GPUs can come in place saving synchronization between them.
 
@@
 Further the processing unit that is chosen has a strong influence on the quality and speed the final network that lies within acceptable error limits will be reached.
-Based on the choice of this other factors come into consideration influencing the process like communication between processing units and parallelism of data between them.
+Based on the choice of this other factors came into consideration influencing the process, like communication between processing units and parallelism of data between them.
 They are not quite as influential as the first two but can lead to a strong improvement or worsening.
@@
 To ease this technology like GPUDirect\cite{nvidiagpudirect2017} and data prefetching\cite{yang2010gpgpu} can be used.
-It may be possible to decrease the effect of the data flow further by introducing a direct memory transfer unit which can indepently organize data and can handle parallel transfers to and from multiple units.
-This could decrease the calculation time to a bare minimum but would requires additional hardware and maybe modification of the used GPU and CPU.
-Although this would break security of the system on broder range because of the unauthorized memory access and changes, it could offer a steep improvement for large calculations on distributed systems.
-
-TODO: Takeaway, different ideas, innnovative
+It may be possible to decrease the effect of the data flow further by introducing a direct memory transfer unit which can indepently organize synchronization and can handle parallel transfers to and from multiple GPUs.
+This could decrease the calculation time to a bare minimum but would require additional hardware and maybe modification of the used GPUs.
+Although this could break some security limits because of unauthorized memory accesses.
 

### 2018-01-10 paper.tex
add something challenges and conclusion

@@
 \section{Choice of Processing Units}
-The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
+The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU, Multi-GPU and GPU-Clustering.
 \subsection{CPU}
@@
 
+%memory limitation
+Another issue appears while training if the training data is too large to fit in memory.
+Then time expensive disk accesses to either very slow HDDs or faster but still relatively slow SSDs are required.
+For example a fast available SSD SAMSUNG 960 PRO can reach up to 3.5 GB/s which is still relatively slow compared to the maximum 31.5 GB/s possible through an PCIe x16 interface.
+To further increase speed a RAID 0 setup can be created consisting of multiple M.2 SSDs with the potential of parallel access and data transfer reaching for instance with 5 960 PRO that could transfer at maximum of around 17 GB/s.
+
+%bandwith between clusters
+This data transfer becomes only more and more concerning if multiple GPU-Cluster are used, because the connection between them may slower than the interfaces of RAM, GPU and storage used therefore creating a new bottleneck.
+This problem can be reduce if training data is stored by the cluster, which has the downside of further increasing the cost of these setups.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CONCLUSION
@@
 They are not quite as influential as the first two but can lead to a strong improvement or worsening.
+We observed that even in example systems that handle communication efficiently about 35 \% of the calculation time is sacrificed to it.
 

### 2018-01-10 paper.tex
current version

@@
 %Performance Factors for Large Scale Machine Learning Applications
-A Beginners Guide to Perfomance Factors of Machine Learning Applications with Focus on Neural Nets and Deep Learning
+Perfomance Factors of Deep Learning Applications: A Beginners Guide
 }
@@
 \maketitle
-
+2
 % As a general rule, do not put math, special symbols or citations
@@
 Afterwards we chose to have a look, on one hand, at small scale neural networks running on a single machine without large data sets and a low query frequency after the initial training, and at the other hand, large scale neural networks operating with more layers and nodes, large data sets and distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
-We divided our survey into first the choice of neural network afterwards choice of processing units, including basic data splitting strategies for multi processing units systems, followed by challenges occuring when using large and small scale applications.
+We divided our survey into first the choice of neural network afterwards choice of processing units, including basic data splitting strategies for multi processing units systems, followed by challenges occuring while using large and small scale applications.
 
@@
 
+%simple
 Most commonly used Neural Networks of the more classic single or non hidden layer kind are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a simple layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
@@
 
+%deep
 A more advanced architecture is to increase the number of hidden layers to enlarge the application possibility of these networks for more complex models these are then called Deep Networks.
@@
 
+%%%%BEGIN PROCESSING UNIT
+
+The next major performance influencing factor will be the choice of the actual processing unit.
+This decision is the most computation time influencing and can also influence the quality of the result.
+We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but posses a smaller virtualization limit than GPUs because of their differentiating architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters,being simple a seperate unit consisting of multiple GPUs, like they are used in supercomputers.
+
+%%%%END PROCESSIONG UNIT
+
 %%%%BEGIN INFORMATION ACTUAL USED IMPLEMENTATIONS
@@
 
-%%%%BEGIN PROCESSING UNIT
-
-The next major performance influencing factor will be the choice of the actual processing unit.
-This decision is the most computation time influencing and can also influence the quality of the result.
-We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but posses a smaller virtualization limit than GPUs because of their differentiating architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters,being simple a seperate unit consisting of multiple GPUs, like they are used in supercomputers.
-
-%%%%END PROCESSIONG UNIT
-
 %%%%BEGIN RELATED WORK
 
-The topic of performance of neural networks and influencing factors has been touched by many benchmarking survey like \cite{shi2016benchmarking} and \cite{qi2016paleo}.	 But we hope to offer a collected view on the most influencing factors in a more abstract style.
+The topic of performance of neural networks and influencing factors has been touched by many benchmarking survey like in "Benchmarking state-of-the-art deep learning software tools" \cite{shi2016benchmarking} and "Paleo: A performance model for deep neural networks" \cite{qi2016paleo}.	 But we hope to offer a collected view on the most influencing factors in a more abstract style.
 
@@
 
+%generation
 If the task includes the generation of data to be output a GAN, VAE or RNN/LSTM.
@@
 
-If the task includes classification or interpretation of visual data neural networks like CNN or DBN can be used.
-For once CNN are a straightforward choice since they are able to compress data and therefore abstract smaller data like the content of an image from the original image.
-But DBN offer also an possibility to complete the task with the major adavantage of being able to first roughly guess values from the training data and then train unsupervised in the fine-tuning phase as shown by Zhong 2016\cite{zhong2016diversified}.
+%classification
+If the task includes classification or interpretation of visual data deep neural networks, like CNN or DBN, can be used.
+For once CNNs are a straightforward choice since they are able to compress data and therefore abstract smaller data like the content of an image from the original image.
+But DBN also offer a possibility to complete the task with the major advantage of being able to first roughly guess values based on the training data and then train unsupervised in the fine-tuning phase as shown by Zhong 2016\cite{zhong2016diversified}.
 
+%sequential data
 If the task includes the classification or intepretation of incoming data over time the ability of RNN and LSTMs to data that is of sequential nature comes in handy.
@@
 The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU, Multi-GPU and GPU-Clustering.
+
 \subsection{CPU}
 As the first most naive option we have a look on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
+
 \subsubsection{Single-thread}
@@
 But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Table I).
+
 \subsubsection{Multi-thread}
@@
 The effect  continues to grow by using more threads but reaches it limits when using more threads than physical cores available.
+
 \begin{table} 
@@
 \end{table}
+
+%limit
 Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
+
 \subsection{GPU}
@@
 
+%restrictions
 Problems lie in the tightly restricted memory of graphic cards that may not be able to contain all nodes with their corresponding weights as well as training dataset.
@@
 
+%speedup numbers
 A main advantage of Multi-GPU is of course the increased computation power available during the training phase, reducing training time by 35\% when doubling the number of GPUs in CNTK and MXNET, 28 \% in Caffe and 10 \% in  Torch and Tensorflow\cite{shi2016benchmarking}.
 
+%memory
 Another advantage is the greater availability of memory somewhat easing the restrictions of the memory restricted GPUs structure.
@@
 
+%speedup figure
 Fig.
@@
 
+%model choice
 This leads to the focus lying on the design to find the best fit of complexity to fit the desired model. Otherwise longer training time or less accuracy of the resulting network will conclude.
 
+%rest trivial
 Further small scale applications are relatively free of performance related complications, because of their relative simplicity and fast modern computer hardware enabling operations which are not optimal to be resolved in a non less desirable time.

### 2018-01-11 paper.tex
add small history and chart

@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[conference]{IEEEtran}
@@
 \usepackage[T1]{fontenc}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
 
@@
 \usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
 
@@
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
 
@@
 \usepackage{graphicx}
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
 
@@
 %Performance Factors for Large Scale Machine Learning Applications
-Perfomance Factors of Deep Learning Applications: A Beginners Guide
+Perfomance Factors of Deep Learning and Shallow Network Applications: A Beginners Guide
 }
@@
 \maketitle
-2
+
 % As a general rule, do not put math, special symbols or citations
@@
  At first we attend the choice of neural network and the impact of this decision.
-Afterwards we chose to have a look, on one hand, at small scale neural networks running on a single machine without large data sets and a low query frequency after the initial training, and at the other hand, large scale neural networks operating with more layers and nodes, large data sets and distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
-We divided our survey into first the choice of neural network afterwards choice of processing units, including basic data splitting strategies for multi processing units systems, followed by challenges occuring while using large and small scale applications.
+Afterwards we chose to have a look, on one hand, at small scale neural networks running on a single machine without large data sets and at the other hand, large scale neural networks operating with more layers and nodes, large data sets and distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
+We divided our survey into first the choice of neural network afterwards choice of processing units, including basic data distribution strategies for multi processing units systems, followed by challenges occuring while using large and small scale applications.
 
@@
 \section{Background and Related Work}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{pop.png}
+\caption{Popularity in \% of the term "deep learning" on Google in relation to maximum peak, plotted with the Google Trends tool provided by Google}
+\end{figure}
+
+%%%%BEGIN DEEP LEARNING
+
+The term "Deep Learning" itself is just a rebranding of the since the 80s existing concept of multilayered perceptrons.
+Its basic idea is to simply use more than 1 layer in a neural network to increase the accuracy of the learned model.
+A good defintion of Deep Learning is delivered by Deng in his 2014 book "Deep learning: methods and applications"\cite{deng2014deep}: "A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification".
+
+Since then Deep Learning has experienced an almost exponential growth in popularity as seen in Figure 1, where the popularity steeply increases around 2014.
+%%%%END DEEP LEARNING
 
@@
 %simple
-Most commonly used Neural Networks of the more classic single or non hidden layer kind are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a simple layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
+Most commonly used shallow Neural Networks are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a single layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
 
@@
 This decision is the most computation time influencing and can also influence the quality of the result.
-We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but posses a smaller virtualization limit than GPUs because of their differentiating architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters,being simple a seperate unit consisting of multiple GPUs, like they are used in supercomputers.
+We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but posses a smaller virtualization limit than GPUs because of their differentiating architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters,being simply a seperate unit consisting of multiple GPUs, like they are used in supercomputers.
 
@@
 
+%Transition
+Next will follow the choice of neural network and its influence on performance.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CHOICE OF NETWORK
 
-\section{Choice of Neural Network}
+\section{Performance impacting choices}
 
@@
 This decision is heaviliy reliant on the application that is to be implemented.
-This decision influences all further actions and has to be carefully made, to assist this decision we will offer some generalization when to use simple neural networks or deep neural network and what specializations of certain neural networks are giving them advantage while doing specific tasks.
+This decision influences all further actions and has to be carefully made, to assist this decision we will offer some generalization when to use shallow neural networks or deep neural network and what specializations of certain neural networks are giving them advantage while doing specific tasks.
 %%%%END IMPORTANCE
@@
 %%%%BEGIN ADVANTAGES DISADVANTAGES SIMPLE AND DEEP
-\subsection{What kind of networks to use}
-\subsubsection{Simple Networks}
+\subsection{Choice of Neural Network}
+\subsubsection{Shallow Networks}
 Simple task can be done by a network which is not deep like a FFN or a RBM.
@@
 But there small size can lead that accuracy decreses when the model becomes more difficult than anticipated.
-Small scale applications often incur that a simple neural network is used because of the smaller training set and shorter computation time.
+Small scale applications often incur that a shallow neural network is used because of the smaller training set and shorter computation time.
 
 \subsubsection{Deep Networks}
-More complex tasks more often require a deep neural network that can learn a more complex model than simple networks and offer a greater accuracy while doing this.
+More complex tasks more often require a deep neural network that can learn a more complex model than shallow networks and offer a greater accuracy while doing this.
 The major downside of these networks that they require a larger amount of time to train and can be oversized for the actual to be represented model, and therefore adding unnecessary computation time.
@@
 We can not offer a clear answer when to use which kind of network, every problem must be individually analyzed and based on the information, and complexity, as well as the acceptable error rate of the final network the decision has to be made.
-For cases that are unclear a simple network can first be constructed,because of their short training time, and if the result is unsatisfying a more complex network can be considered.
+For cases that are unclear a shallow network can first be constructed,because of their short training time, and if the result is unsatisfying a more complex network can be considered.
 Also it is advisable to use a deep network if the data is either of image or audio nature and/or needs to include changes over time of an input signal\cite{PattersonGibson17}.
@@
 %%%%DETAILS SIMPLE AKA SMALL
-\subsection{Simple network Nodes}
-If a simple network shall be used the next choice will be the amount of nodes in the hidden layer.
+\subsection{Shallow network node amount}
+If a shallow network shall be used the next choice will be the amount of nodes in the hidden layer.
 This is of course dependent one foremost the number of inputs and outputs, the number of hidden nodes should deviate to strong from them.
 How strong this deviation really is is dependent on the chosen neural network, for example a feed forward network profits when the number is equal or larger than the number of inputs, while an auto encoder probably has fewer hidden nodes.
-For a simple network this decision can be if too far from the optimum less fatal than for a deep network, while a simple network may lose some accuracy or gain additional training time a deep network can experience these effects at a greater scale.
+For a shallow network this decision can be if too far from the optimum less fatal than for a deep network, while a shallow network may lose some accuracy or gain additional training time a deep network can experience these effects at a greater scale.
 
@@
 %%%%DETAILS DEEP AKA LARGE AND SMALL
-\subsection{Deep networks Nodes}
-Like in simple networks the amount of nodes to be used is an important value that determines the final ability of the network to fulfill its task in the required accuracy and speed.
+\subsection{Deep networks node amount}
+Like in shallow networks the amount of nodes to be used is an important value that determines the final ability of the network to fulfill its task in the required accuracy and speed.
 The acutal number of nodes chosen at the end is of course a bit more complex in deep networks for example networks like DCN require a special node configuration based on their basic idea(step by step decreasing of nodes per layer until desired compression reached) , or DN(opposite of DCN).
@@
 
+%Transition
+To further optimize training time we will next touch the topic of processing units which can be used.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%BEGIN CHOICE OF PROCESSING UNIT
 
-\section{Choice of Processing Units}
+\subsection{Choice of Processing Units}
 The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU, Multi-GPU and GPU-Clustering.
@@
 %speedup figure
-Fig.
-1 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs used.
+Figure 2 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs used.
 Although costly this leads to a constant improvement.
@@
 \subsubsection{Synchronization between GPUs}
-For synchronizing Multi-GPU setups like in Figure 3 there are theoretically two choices.
+For synchronizing Multi-GPU setups like in Figure 4 there are theoretically two choices.
 One being the synchronous method in which after a value is calculated they are shared with every other GPU.
@@
 In doing this we again increase raw computation power, but also increase the total amount of communication required.
-Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in (Fig.
-2)
+Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in (Figure 3)
 
@@
 This creates the best performance on large configurations with many clusters\cite{wang2016deep}.
+Which challenges this and other characteristics of this incur will be explained in the next section.
 
@@
 %%Data flow of training data
-The data flow is a problem on multiple fronts.
+The \emph{data flow} is a problem on multiple fronts.
 For one the transfer of the training data from main memory to GPU memory can be the first bottleneck restricting faster processing speeds from being realized.
@@
 %%data flow synchronization
-Additional data flow is created when synchronizing GPUs with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations.
+Additional data flow is created when \emph{synchronizing GPUs} with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations.
 But both create an communication overhead because during transmission of values the GPU cannot operate on the neural network training.
@@
 Also more GPUs lead to more communication if synchronization is required after very step, so they increase the amount of data managed by the CPU or communication server.
-This applies to configuration like Figure 3 and Figure 4.
+This applies to configuration like Figure 4 and Figure 5.
 
 %asynchronous
-If an asynchronous training approach is used problems may occur influencing the result, if updates are to sporadic and different versions of values drift too far apart.
+If an \emph{asynchronous training} approach is used problems may occur influencing the result, if updates are to sporadic and different versions of values drift too far apart.
 To avoid this problem GPU-Cluster configurations do not rely on complete asynchronous training startegies, but on the already described mixed-asynchronous minimizing this problem while still reducing possible communication overhead.
 
-To ease the effect of GPU to GPU communication,by relieveing the CPU of managing them, technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead.
+To ease the effect of \emph{GPU to GPU communication},by relieveing the CPU of managing them, technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead.
 Also shared memory between GPUs can come in place saving synchronization between them.
@@
 %memory limitation
-Another issue appears while training if the training data is too large to fit in memory.
+\emph{Memory limitation} is another issue appears while training if the training data is too large to fit in memory.
 Then time expensive disk accesses to either very slow HDDs or faster but still relatively slow SSDs are required.
@@
 %bandwith between clusters
-This data transfer becomes only more and more concerning if multiple GPU-Cluster are used, because the connection between them may slower than the interfaces of RAM, GPU and storage used therefore creating a new bottleneck.
+\emph{Cluster communication} becomes only more and more concerning if more GPU-Clusters are used, because the connection between them may slower than the interfaces of RAM, GPU and storage used therefore creating a new bottleneck.
 This problem can be reduce if training data is stored by the cluster, which has the downside of further increasing the cost of these setups.

### 2018-01-11 paper.tex
add suggested changes

@@
 Machine Learning is probably the most present topic of computer science in general media.
-This popularity only increased over the recent years thorugh widely publicly known applications ranging from DeepBlue(not so recent) to IBM Watson and AlphaGo.
-Because of these successes Machine Learning is sometimes viewed as omnipotent solution to every imaginable problem.
-This is most certainly not the case, Neural Nets and alike are powerful but require a large amount of design and implementation work.
-We want to give a short introduction to performance influencing factorts of neural networks during training and the resulting network.
+This popularity only increased over the recent years thorugh widely publicly known applications ranging from DeepBlue to IBM Watson and AlphaGo.
+Because of these successes Machine Learning is sometimes viewed as an omnipotent solution to every imaginable problem.
+This is most certainly not the case, based on our studies we can conclude that their capabilities depend on done design and engineering work.
+In this paper we want to give an short introduction into the most performance influencing factors to performance, during the most time intensive phase, the training.
 
@@
 Machine Learning seems to be everywhere around us.
-With the recent success of Neural Networks, and other Machine Learning tools, especially the, in general media, well reported learners like AlphaGo, Silver or IBM Watson, these are experiencing a kind of renaissance their usage has spread from industrial(e.g.
-Self-driving cars rely on autonomous interpretation of camera feed and sensor output that can be optimized by using machine learning) to retail.
-But also in modern automation and research, machine learning becomes a greater field managing to find more efficient ways to achieve better results in many expertise.
+With the recent success of Neural Networks, and other Machine Learning tools, especially  in general media well reported learners like AlphaGo or IBM Watson, these technologies are experiencing a kind of renaissance.
+Their usage has spread from industrial(e.g.
+self-driving cars, which rely on autonomous interpretation of camera feed and sensor output using machine learning) to retail.
+Also in modern automation and research, machine learning becomes a greater tool for managing to find more efficient ways to achieve better results in many cases.
 
-In this paper we want to observe and identify the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scale applications.
+In this paper we want to observe and identify the most performance influencing factors of these networks while designing and implementing them for specific applications, with both small and large scale deployments.
 We expect that this study will provide help for readers interested in learning about the the initial design of neural networks, requirements for managing data, and operations of an efficient neural network.
 
-TODO: related surveys
+During our study we found no comprehensive study touching this thema not being based on the comparison of frameworks or on an experimental execution on different platforms or frameworks.
+This paper is based on these publications.
 
- At first we attend the choice of neural network and the impact of this decision.
-Afterwards we chose to have a look, on one hand, at small scale neural networks running on a single machine without large data sets and at the other hand, large scale neural networks operating with more layers and nodes, large data sets and distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
-We divided our survey into first the choice of neural network afterwards choice of processing units, including basic data distribution strategies for multi processing units systems, followed by challenges occuring while using large and small scale applications.
+ At first we attend to the choice of neural network and the impact of this decision.
+Afterwards we have a look, on one hand, at small scale neural networks running on a single machine without large data sets and on the other hand, large scale neural networks operating with more layers and nodes, large data sets and a distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
+We divide our survey into, the choice of neural network, the choice of processing units, including basic data distribution strategies for multi-processing, a discussion of the challenges in running small and large scales.
+
+%%%%BEGIN RELATED WORK
+
+The topic of considering the performance of neural networks and the influencing factors, has been covered in many studies like
+ in "Benchmarking state-of-the-art deep learning software tools" \cite{shi2016benchmarking} and "Paleo: A performance model for deep neural networks" \cite{qi2016paleo}.
+In our study we hope to offer a collected view, which could serve as an entry point for developers seeking to work with these tools.
+%%%%END RELATED WORK
 
@@
 \section{Background and Related Work}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{neuralnetworks.png}
+\caption{"A mostly complete chart of Neural Networks" by "The Asimov Institute" extract of only in this paper referenced networks\cite{zoo}}
+\end{figure}
 
@@
 
+In this section we introduce some commonly used models, we consider some performance factors and we refer readers to related work in evaluating performance factors of these models.
+For this we begin by showing some of the history of Deep Learning followed by clarification of the distinction between what we define as small and large scale networks.
+
 The term "Deep Learning" itself is just a rebranding of the since the 80s existing concept of multilayered perceptrons.
-Its basic idea is to simply use more than 1 layer in a neural network to increase the accuracy of the learned model.
+Its basic idea is to simply use more than 1 layer in a neural network to increase the accuracy of the learned model\cite{deeplearning101}.
 A good defintion of Deep Learning is delivered by Deng in his 2014 book "Deep learning: methods and applications"\cite{deng2014deep}: "A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification".
+Additionally to the reuse of multilayer perceptrons the ideas of Deep Belief Networks and unsupervised pretraining have been introduced.
 
@@
 
-As a small scale neural network applications we define a network with few neural nodes inside the network and because of this rather limited size it is most of the time run on a single machine and or single processing unit which can complete the training process of the net in a reasonable time.
+As a small scale neural network applications we define a network with few and because of this rather limited size it. Most of the time such a network is run on a single machine.
 
@@
 
-On the other hand as a large scale neural network applications we define a network with a large number of nodes which can because of the complexity of the resulting connections take a large amount of time to complete training if computed on one processing unit and therefore will take use of a distributed system consisting of multiple units in a single system or more servers to synchronize multiple systems of processing units and distribute the required data.
-
+On the other hand as a large scale neural network applications we define a network with a large number of nodes which can, because of the complexity of the resulting connections take a large amount of time to complete training if computed on one processing unit.
+Therefore it requires a system capable of running the training parallel in a single processing unit, a system containing multiple processors in one node, or a multinode system.
+Or in other words a system capable of parallelization, since this is of highest concern in large scale networks.
 %%%%END  LARGE SCALE NEURAL NET DESCRIPTION
@@
 %simple
-Most commonly used shallow Neural Networks are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a single layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
+The most commonly used shallow Neural Networks are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a single layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
 
@@
 
-Auto Encoders(AE) are of an similar build but the hidden layer compresses the input information into fewer nodes while these are later again decompressed to output nodes that match the previously given input nodes.
+Auto Encoders(AE) are of an similar build but the hidden layer compresses the input information into fewer nodes while these are later again decompressed to output nodes that match the previously given input nodes, therefore AE can be classified as a generative network.
 
@@
 A more advanced architecture is to increase the number of hidden layers to enlarge the application possibility of these networks for more complex models these are then called Deep Networks.
-The most commonly used ones are Variational Auto Encoders , Deep Belief Networks, Generative Adversial Networks, Recurrent Neural Networks ,e.g.
+The more commonly used ones are Variational Auto Encoders , Deep Belief Networks, Generative Adversial Networks, Recurrent Neural Networks ,e.g.
 Long short-term memory networks, Deep Convolutional Neural Networks, Deconvolutional Networks, and Recursive Neural Networks.
 
-Variational Auto Encoders(VAE) consists of a encoder, sampler and decoder and are able to to generate output based on specification given to the VAE.
+Variational Auto Encoders(VAE) consists of an encoder, sampler and decoder and are able to to generate output based on specification given to the VAE.
 
@@
 data and the FFN to adjust the learned values slightly to improve the result.
+The pretraining is done in an unsupervised manner enabling the network to learn structural features of the input in each layer.
+These networks are at the core of the renaissance of neural networks, stemming from the work of Hinton 2006\cite{hinton2006fast}.
 
-Generative Adversial Networks(GAN) like the VAE creates random output based on the model found in the given training data.
+Generative Adversial Networks(GAN) like the VAE create random output based on the model found in the given training data.
 VAE are a specialization of GAN because of the restrictions that can be set for the output which is not possible in a GAN.
 
-Deep Convolutional Networks(DCN) are based on Convolutional Nodes compressing the incoming information into the needed base model and then calculate the required output.\cite{PattersonGibson17} They are used to abstract input information into more general models.
+Deep Convolutional Networks(DCN) are based on Convolutional Nodes compressing the incoming information into the needed base model and then calculating the required output.\cite{PattersonGibson17} They are used to abstract input information into more general models.
+These networks especially are being used to process images.
 
@@
 
-Long short-term memory(LSTM)\cite{hochreiter1997long} are a variant of Recurrent Neural Networks(RNN) including additionally memory cells and gates that control the content of the cells.
+Long short-term memory(LSTM)\cite{hochreiter1997long} are a variant of Recurrent Neural Networks(RNN) including additionally memory cells and gates that control the content of the cells, that are able to remember patterns in sequential data.
 This model has been able to improve learning process of RNN and results.
@@
 The reason why the training of neural networks is so expensive can be traced back to the algorithm used to train all commonly used networks, which is backpropagation.
-Backpropagation is based on the idea that we calculate the deviation of the output values from the actual result and then use this backwards thorugh the layers of the net to calculate a correction,based on the error of the node and the learning rate, for each connection weight until the input layer is reached\cite{riedmiller1993direct}.
+Backpropagation is based on the idea that we first calculate the deviation of the output values from the actual result and then use this backwards thorugh the layers of the net to calculate a correction,based on the error of the node and the learning rate, for each connection weight until the input layer is reached\cite{riedmiller1993direct}.
 Some Networks like DBN use a slightly
-diverged "Gentle" Backpropagation, which uses a lowered learning rate after the initial values have been found from the training data with the help of the pretrain RBM \cite{PattersonGibson17}.
+diverged "Gentle" Backpropagation, which relies on a learning rate after the initial values have been found from the training data with the help of the pretrained RBM \cite{PattersonGibson17}.
 %%%%END TRAINING ALGORITHMS
@@
 
-Another performance influencing factor besides the actual neural net and the used learning algorithm is the amount of data required for the net to be trained and the size of the input of each training example.While this is mostly important for large scale applications it can also be hindering for smaller ones.
+Another performance influencing factor besides the actual neural net and the learning algorithm, is the amount of data required for the net to be trained and the size of each training example.While this is mostly important for large scale applications, it can also be hindering for smaller ones.
 For example if we use a Convolutional Neural Network to classify the content of a greyscale picture we get for small $512*512$ sized images $262144$ inputs that each need to be transferred to the processing unit, CPU or GPU, for each example.
 
-This large amount of data can for one take some time to be transferred, if for example this data package is required by an external processing unit like a GPU other units can not  receive new data and therefore lose performance due to them being idle.
+This large amount of data can, for one, take some time to be transferred, if for example this data package is required by an external processing unit like a GPU other units cannot receive new data and therefore the performance drops due to them being idle.
 Also if too much data is needed in relation to the actual output the data storage can achieve enduring performance loss will occur.
@@
 
-Additionally the synchronization can be somewhat difficult and extensive for large neural networks on multiple processing units because of the interchange of new weights calculated between them after a certain number of iterations on their training cases.
+Furthermore if the training data exceeds the memory for the processing unit, the problem of delivering data to all units becomes more pressing.
 
@@
 The next major performance influencing factor will be the choice of the actual processing unit.
-This decision is the most computation time influencing and can also influence the quality of the result.
-We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but posses a smaller virtualization limit than GPUs because of their differentiating architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters,being simply a seperate unit consisting of multiple GPUs, like they are used in supercomputers.
+This decision is the most computation time influencing.
+This decision can have the biggest influence on computation time.
+We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but possess a smaller parallelization limitation than GPUs because of their different architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters, being simply a group of GPUs either in a single node or spanning multiple nodes, like they are used in supercomputers.
 
@@
 
-The most common neural network frameworks which we, because of their wide popularity, will mostly view are Tensorflow\cite{abadi2016tensorflow}, Caffe\cite{jia2014caffe}, Torch\cite{collobert2002torch}, CNTK\cite{gitcntk}, deeplearning4j\cite{websitedl4j}, Keras\cite{websitekeras}, MXNET\cite{websitemxnet}.
-These frameworks have been benchmarked in numerous settings and applications and can help us show performance influencing factors.
+The most common neural network frameworks which we, because of their wide popularity, have found in performance studies are Tensorflow\cite{abadi2016tensorflow}, Caffe\cite{jia2014caffe}, Torch\cite{collobert2002torch}, CNTK\cite{gitcntk}, deeplearning4j\cite{websitedl4j}, Keras\cite{websitekeras}, and MXNET\cite{websitemxnet}.
+
+% These frameworks have been benchmarked in numerous settings and applications and can help us show performance influencing factors.
 
@@
 %%%%END INFORMATION ACUTAL USED IMPLEMENTATIONS
-
-%%%%BEGIN RELATED WORK
-
-The topic of performance of neural networks and influencing factors has been touched by many benchmarking survey like in "Benchmarking state-of-the-art deep learning software tools" \cite{shi2016benchmarking} and "Paleo: A performance model for deep neural networks" \cite{qi2016paleo}.	 But we hope to offer a collected view on the most influencing factors in a more abstract style.
-
-%%%%END RELATED WORK
 
@@
 This decision is heaviliy reliant on the application that is to be implemented.
-This decision influences all further actions and has to be carefully made, to assist this decision we will offer some generalization when to use shallow neural networks or deep neural network and what specializations of certain neural networks are giving them advantage while doing specific tasks.
+Such decision influences all further actions and has to be carefully made.
+
+To assist this decision we offer some general guidelines for when to use shallow vs. deep networks, and what specializations of certain networks could be advantageous for specific tasks.
 %%%%END IMPORTANCE
@@
 \subsubsection{Shallow Networks}
-Simple task can be done by a network which is not deep like a FFN or a RBM.
+Simple task can be done by a network which is not deep, like a FFN or a RBM.
 They are faster to train and even larger training sets can be completed in a reasonable time without multithreading or usage of GPUs.
-But there small size can lead that accuracy decreses when the model becomes more difficult than anticipated.
-Small scale applications often incur that a shallow neural network is used because of the smaller training set and shorter computation time.
+But their limited depth can decrease the accuracy of the model when the learned task is more complex than expected.
+In other words, these models have a risk to overfit the training data.
+Accordingly, to use them, careful validation is required.
 
 \subsubsection{Deep Networks}
-More complex tasks more often require a deep neural network that can learn a more complex model than shallow networks and offer a greater accuracy while doing this.
-The major downside of these networks that they require a larger amount of time to train and can be oversized for the actual to be represented model, and therefore adding unnecessary computation time.
+More complex tasks often require a deep neural network that can learn a more complex model than shallow networks and offer a greater accuracy while doing this.
+The major downside of these networks is that they require a larger amount of time to train and can be oversized for the actual learning task and might adding unnecessary computation time.
 Large scale applications mostly contain deep networks because of their larger training sets and more complex models underlying the actual task.
 
-We can not offer a clear answer when to use which kind of network, every problem must be individually analyzed and based on the information, and complexity, as well as the acceptable error rate of the final network the decision has to be made.
+We can not offer a clear answer about when to use which kind of network, every problem must be individually analyzed and based on the information, and complexity, as well as the acceptable error rate of the final network the decision has to be made.
 For cases that are unclear a shallow network can first be constructed,because of their short training time, and if the result is unsatisfying a more complex network can be considered.
-Also it is advisable to use a deep network if the data is either of image or audio nature and/or needs to include changes over time of an input signal\cite{PattersonGibson17}.
+Also it is advisable to use a deep network if the learning task involves image or audio understanding, or the task needs to include recurring changes over time of an input signal\cite{PattersonGibson17}.
 %%%%END ADVANTAGES DISADVANTAGES SIMPLE AND DEEP
@@
 %%%%END SPECIFIC NETS
+
+Finally, if the task has been covered in the literature, it is worthwhile to use existing publically available networks like AlexNet, GoogleNet and others.
 

### 2018-01-12 paper.tex
added some changes to the last 4 pages

@@
 %Performance Factors for Large Scale Machine Learning Applications
-Perfomance Factors of Deep Learning and Shallow Network Applications: A Beginners Guide
+Perfomance Factors of Deep Learning and Shallow Neural Network Applications: A Beginners Guide
 }
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{pop.png}
-\caption{Popularity in \% of the term "deep learning" on Google in relation to maximum peak, plotted with the Google Trends tool provided by Google}
-\end{figure}
-
 %%%%BEGIN DEEP LEARNING
@@
 
-Since then Deep Learning has experienced an almost exponential growth in popularity as seen in Figure 1, where the popularity steeply increases around 2014.
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{pop.png}
+\caption{Popularity in \% of the term "deep learning" on Google in relation to maximum peak, plotted with the Google Trends tool provided by Google}
+\end{figure}
+
+Since then Deep Learning has experienced an almost exponential growth in popularity as seen in Figure 2, where the popularity steeply increases around 2014.
 %%%%END DEEP LEARNING
@@
 If a shallow network shall be used the next choice will be the amount of nodes in the hidden layer.
-This is of course dependent one foremost the number of inputs and outputs, the number of hidden nodes should deviate to strong from them.
-How strong this deviation really is is dependent on the chosen neural network, for example a feed forward network profits when the number is equal or larger than the number of inputs, while an auto encoder probably has fewer hidden nodes.
-For a shallow network this decision can be if too far from the optimum less fatal than for a deep network, while a shallow network may lose some accuracy or gain additional training time a deep network can experience these effects at a greater scale.
+This is of course dependent on the number of inputs and outputs.
+The number of hidden notes should not deviate too strongly from these two.
+The amount of the deviation is really dependent on the chosen neural network, for example a feed forward network profits when the number is equal or larger than the number of inputs, while an auto encoder probably has fewer hidden nodes.
+When the number of nodes is higher than the optimum, there can be a loss of accuracy or an increase in the number of epochs needed for training.
+In general this effect can be larger for deep networks than for shallow ones.
 
@@
 Like in shallow networks the amount of nodes to be used is an important value that determines the final ability of the network to fulfill its task in the required accuracy and speed.
-The acutal number of nodes chosen at the end is of course a bit more complex in deep networks for example networks like DCN require a special node configuration based on their basic idea(step by step decreasing of nodes per layer until desired compression reached) , or DN(opposite of DCN).
-So how the node configuration looks like is strongly dependent on the net chosen.
+The actual number of nodes chosen at the end is of course a bit more complex in deep networks.
+For example networks like DCN require a special node configuration based on their basic idea(step by step decreasing of nodes per layer until desired compression reached) , or DN(opposite of DCN).
+So how the node configuration looks like is strongly dependent on the net chosen and the training task.
 
@@
 \subsection{Specific deep networks}
-Based on the content of the task a few networks can be more fitting than others to fulfill it.
-To deliver a shot overview of it we use a classification by Patterson\cite{PattersonGibson17} for the basis of our advise.
+Based on the content of the task a few networks can be more fitting than others to fulfill it. To deliver a short overview of it we use a classification by Patterson\cite{PattersonGibson17}.
 
 %generation
-If the task includes the generation of data to be output a GAN, VAE or RNN/LSTM.
+If the task includes the generation of data a GAN, VAE or RNN/LSTM is recommended.
 Which of these to choose is strongly dependent of the specific parameters of the task.
-For example if we want to generate random new data following the model of our training data the closest to this would be a GAN because it is by design designated to fulfill such tasks.
+For example if we want to generate random new data following the model of our training data, the closest to this would be a GAN because it is by design built to fulfill such tasks.
 But if new data which is not completely random is to be generated a VAE can be considered which offers by design this possibility.
-In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} (corresponding paper \cite{zhu2016generative}) or text generation with LSTM\cite{gittesttensorflow} (paper describing base approach\cite{sutskever2011generating}).
+In practice there are lots of generation examples available using these networks for example image generation with GAN\cite{junyanz2017} (For the interested reader, we suggest \cite{zhu2016generative} as a paper describing this approach) or text generation with LSTM\cite{gittesttensorflow} (Described in \cite{sutskever2011generating}).
 
@@
 %sequential data
-If the task includes the classification or intepretation of incoming data over time the ability of RNN and LSTMs to data that is of sequential nature comes in handy.
+If the task includes the classification or intepretation of incoming data over time the ability of RNN and LSTMs to remember data comes in handy.
 For example voice recognition is possible to implement with the help of LSTMs as shown by Soltau 2016\cite{soltau2016neural}.
@@
 \subsection{Choice of Processing Units}
-The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU, Multi-GPU and GPU-Clustering.
+The next important choice is the choice of the processing units.
+We distinguish them into three main models, CPU, GPU, Multi-GPU and GPU-Clustering.
+In this section we discuss how these units can be used for deep learning.
+We base our discussion on published results of experiments \cite{shi2016benchmarking} and \cite{sastre2017scalability}.
 
 \subsection{CPU}
-As the first most naive option we have a look on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
+As the first most naive option we have a look on the CPU for training the neural net with either a simple single thread, or the more complex multi-thread.
 
 \subsubsection{Single-thread}
-The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required .
-But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Table I).
+The most naive implementation is the single-thread CPU calculation.
+Because of its sequential execution and calculation, it can be the slowest (when not taking into account the use of vectorized instructions).
+For networks with a large number of nodes this approach is simply unfeasible because of the excessive calculation time.
+Since this approach does not utilize the full capacity of the CPU, it can result in performance losses when compared to multi-threaded solutions (Table I).
 
 \subsubsection{Multi-thread}
-A more complex approach consist of utilizing the multi-core characterisitics of modern CPUs, by splitting of training calculation with only requirements to already calculated values.
+A more complex approach consists of utilizing the multi-core characterisitics of modern CPUs, by parallelizing the calculations in the training that only depend on previously calculated values.
 This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm.
-The effect  continues to grow by using more threads but reaches it limits when using more threads than physical cores available.
+As researchers observe, the benefits continue to grows by using more threads but reaches its limits when using more threads than physical cores available.
 
@@
 %limit
-Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
+Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with \emph{hyper-threading} technologies, which can lengthen the required calculation time because of a more inefficient usage of the CPU\cite{shi2016benchmarking}.
 
@@
 A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs.
-For example with the help of  NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}.
-Almost all recent neural net learning frameworks support GPUs because of its excellent performance rating for neural networks.
-
+For example with the help of  NVIDIAs CUDA this can be done efficiently and results in a speed-up of deep learning algorithms of up to 3 times when compared to CPU-based approaches.
+Almost all deep learning frameworks support GPUs.
 %restrictions
-Problems lie in the tightly restricted memory of graphic cards that may not be able to contain all nodes with their corresponding weights as well as training dataset.
-This leads to some communication overhead depending on the communication strategy chosen.
+Problems lie in the tightly restricted memory of graphic cards that may not be able to contain all nodes with their corresponding weights.
+Furthermore there might be some communication overhead between CPUs and GPUs, since the latter can have a higher throughput, and there might be an imbalance making the communication of data and results a bottleneck.
+Accordingly, one solution, the overlapping of computation of communication is an important technique.
 
@@
 %speedup figure
-Figure 2 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs used.
+Figure 3 visualizes this improvement showing the reduction of training time steadily with the increasing of amount of GPUs.
 Although costly this leads to a constant improvement.
-Higher Numbers of GPUs can be again slower or offer no significant improvement if used the same as smaller amounts because of an increasing communication overhead and bottleneck in the CPU, which is responsible for managing seperate GPUs.
+Higher Numbers of GPUs can be again slower or offer no significant improvement if used in a single system because of an increasing communication overhead and bottleneck for the data flow in the CPU, which is responsible for managing separate GPUs.
 
 \subsubsection{Parallelism between GPUs}
-There are two kind of parallelism approaches applicable to the problem, first there is data parallelism which splits the training data into multiple sets, mostly used when the training set is splitable into multiple subsets.
+There are two kind of parallelism approaches applicable to the problem, first there is data parallelism which splits the training data into multiple sets.
+Each subset and a complete replica of the model is assigned to a processing unit, such that each can calculate the gradients for all the layers.
+Finally, before moving to the next epoch, the gradients of all units are aggregates such that the parameters can be updated, shipping a new replica for the next epoch.
 This can be done by dividing the data by case or by feature.
-Second model parallelism in which the net is divided into separately runnable parts which later can be joined together again, this method is fitting for single system distribution in which the GPUs share one PCIe bus, but lead to the result being more vulnerable to calculation failures \cite{sastre2017scalability}.
+The second approach is model parallelism in which the net is divided into separately runnable parts, each part can be assigned to a separate unit such that all units train over the complete data, but each calculates gradients for a subset of neural network nodes.
+Naturally, inter-processing unit communication might be needed when the gradients have to be propagated to other processors.
+For networks that do not have fully connected layers, such that the distribution can be made avoiding (to an extent) inter-unit processing.
+This method is also fitting for single node distributions in which the GPUs share one PCIe bus, but lead to the result being more vulnerable to calculation failures \cite{sastre2017scalability}.
 
@@
 For synchronizing Multi-GPU setups like in Figure 4 there are theoretically two choices.
-One being the synchronous method in which after a value is calculated they are shared with every other GPU.
-This method is applicable to all Multi-GPU configuration and delivers an acceptable result\cite{wang2016deep}.
-The update of values for all GPUs occurs only after the updated values have been send by each of them.
+One being the synchronous method in which after a value is calculated it is immediately shared with every other GPU.
+This method is applicable to all Multi-GPU configurations and delivers an acceptable result\cite{wang2016deep}.
+The update of values for all GPUs occurs only after the updated values have been sent by each of them.
 Until a large amount of over 40 GPUs are used this mehtod can be used without limitation\cite{sastre2017scalability}.
-There is also the possibility to run the GPUs asynchronously, by letting them use different versions of the values and only updating them periodically by sending an update request to the CPU.
-This can lead to a slight improvement of the training speed but is not guaranteed depending on the training model\cite{wang2016deep}.
+There is also the possibility to run the GPUs asynchronously, by letting them use different versions of the values and only updating them periodically by sending an update request.
+This can lead to a slight improvement of the training speed but it might also have a slight detrimental effect in the learning process\cite{wang2016deep}.
+This depends on the network model and training algorithm.
 
@@
 Like the Multi-GPU setup this can be done by creating either data or model parallelism.
-Combination are also possible when using GPU-Clusters, for instance in a setup consisting of multiple Clusters and one single communication server a model parallelism can be used to distribute the model on the different clusters, but they internally use data parallelism to split work\cite{wang2016deep}.
+Combination are also possible, for instance in a setup consisting of multiple Clusters and one single communication server model parallelism can be used to distribute the model on the different clusters, but they internally can use data parallelism to split the work\cite{wang2016deep}.
 
 \subsubsection{Synchronization of multiple GPU-Clusters}
-Similar to parallelism the synchronization of GPU-Clusters can be undertaken either with the synchronous or asynchronous like in Multi-GPU configurations.
-Again these tactics can be combined to create a mixed-asynchronous strategy, using an asynchronus training on inter-clusters level and cluster internal synchronous training\cite{sastre2017scalability}.
+The synchronization of GPU-Clusters can be undertaken either with synchronous or asynchronous approaches.
+These tactics can also be combined to create a mixed-asynchronous strategy, using an asynchronus training on inter-cluster level and on intra-cluster level synchronous training\cite{sastre2017scalability}.
+This approach is specially reasonable considering that intra-cluster communication will usually be much faster than inter-cluster communication.
 This creates the best performance on large configurations with many clusters\cite{wang2016deep}.
@@
 \section{Large scale challenges and Small scale challenges}
-Dependent on the choice of scale each one has its own challenges to be considered during design and implementation of the neural network.
+Different scales of deep learning models have different challenges that need to be considered during the design and implementation of the neural network.
 
 \subsection{Small scale challenges}
-First and foremost it is important to notice that the biggest restrictions of small scale applications is the limited computation power available during the initial training process, later during usage of the net this is less impeding because either the net is completely pretrained and not changed later or the only change is some optimization when the network receives a query.
+First and foremost it is important to notice that the biggest restrictions of small scale applications is the limited computation power available during the initial training process, later during the usage of the net for inference (also called prediction serving), this is less impactful because either the net is completely pretrained and not changed later or the changes are limit like some optimization when the network receives a query.
 
 %model choice
-This leads to the focus lying on the design to find the best fit of complexity to fit the desired model. Otherwise longer training time or less accuracy of the resulting network will conclude.
+This leads to the focus lying on the design to find the best fit of complexity to fit the desired model. Otherwise longer training time or less accuracy of the resulting network will occur.
 
 %rest trivial
-Further small scale applications are relatively free of performance related complications, because of their relative simplicity and fast modern computer hardware enabling operations which are not optimal to be resolved in a non less desirable time.
+Further small scale applications are relatively free of performance related complications, because of their relative simplicity and fast modern computer hardware.
 
 \subsection{Large scale challenges}
-Opposing to small scale applications large scale ones to not have the actual processing power as their highest pending restriction but the communication and data transfer between all acting processing units.
-This problem occurs because a larger data flow has to be managed to more processing units compared to small scale applications.
+In contrast to small scale applications, the most challenging factor for large scale applications is the proper scheduling of operators, to reduce the costly communication and synchronization between processing units.
+This problem occurs because a larger data flow has to be managed over more processing units compared to small scale applications.
 
@@
 %%Data flow of training data
-The \emph{data flow} is a problem on multiple fronts.
-For one the transfer of the training data from main memory to GPU memory can be the first bottleneck restricting faster processing speeds from being realized.
-For example if the GPU only receives data if no calculation is currently active this slows down the training process unnecessarily.
+The \emph{data flow} is a problem on multiple fronts.On one side the transfer of the training data from main memory to GPU memory can be the first bottleneck restricting faster processing speeds from being realized.
+For example if the GPU only receives data and no calculation is currently active this slows down the training process unnecessarily.
 To increase the speed of the calculation data can be fetched during the processing of already present data, this is called data prefetching as described and implemented by Yang 2010 \cite{yang2010gpgpu}.
-This solution can solve the problem partially but reaches it limits if used with multiple GPUs for example configuration (Fig.
-3), which leads that the CPU will become a bottleneck if the training data of each individual GPU is too large to fit in the GPU memory, since the CPU can only manage transfer to the GPUs one at a time.
-For example a GPU equipped with PCI-Express 4.0 with a x16 Throughput can at best receive 31.5 GB/s, if we assume that all training data can be fit into main memory, that way if we need to refresh the complete memory of a modern GPU like NVIDIA GTX 1080, we need 0.253968 s, so to refresh all GPUs memory without including any overhead produced by switching operations between the GPUs we have an absolute transfer time of 1.015873 s, and for 3 units 0.7619 s, so if our GPU completes the desired calculation faster, for example 0.5 s, we have an 0.2619 s period in which our GPU idles every iteration, assumed that we can transfer all data during the calculation time.
+This solution can solve the problem partially but reaches its limits when used with multiple GPUs.
+For example configuration (Fig.
+3), this makes the CPU-GPU transfer the single bottleneck, if the data for an iteration (training data plus the model) in each individual GPU is too large to fit in the GPU memory.
+For example a GPU equipped with PCI-Express 4.0 with a x16 Throughput can at best receive 31.5 GB/s, if we assume that all training data can reside in the CPU main memory, that way if we need to refresh the complete memory of a modern GPU like NVIDIA GTX 1080, we need 0.253968 s, so to refresh all GPUs memory without including any overhead produced by switching operations between the GPUs we have an absolute transfer time of 1.015873 s, and for 3 units 0.7619 s, so if our GPU completes the desired calculation faster, for example 0.5 s, we have a 0.2619 s period in which our GPU is idle every iteration.
 Therefore we have an effective calculation computation time each phase of 65.62 \%.
-This problem only enlargens if the main memory is to small to fit the complete training data since connections to solid state disk are too slow and hard disk do physically do not reach the required throughput.
+This problem only grows if the main memory is too small to fit the complete training data since connections to solid state disk are too slow and hard disk do physically do not reach the required throughput.
 
@@
 Additional data flow is created when \emph{synchronizing GPUs} with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations.
-But both create an communication overhead because during transmission of values the GPU cannot operate on the neural network training.
+Both create an communication overhead because during transmission of values the GPU cannot operate on the neural network training.
 
@@
 %synchronous
-Synchronous training, while being recommend for setups that have 40 GPUs or less, can create idling time if one unit is slower than others, since all updates have to be send to the server before the next iteration can be undertaken this slows down all other units.
-This problem becomes more probable the more units are used.
+Synchronous training, while being recommend for setups that have 40 GPUs or less, can create idle time if one unit is slower than others, since all updates have to be sent to the server before the next iteration can be undertaken, this slows down all other units.
+This problem becomes more likely the more units are used.
 Also more GPUs lead to more communication if synchronization is required after very step, so they increase the amount of data managed by the CPU or communication server.
-This applies to configuration like Figure 4 and Figure 5.
+This applies to configurations like Figure 5 and Figure 6.
 
 %asynchronous
-If an \emph{asynchronous training} approach is used problems may occur influencing the result, if updates are to sporadic and different versions of values drift too far apart.
-To avoid this problem GPU-Cluster configurations do not rely on complete asynchronous training startegies, but on the already described mixed-asynchronous minimizing this problem while still reducing possible communication overhead.
+If an \emph{asynchronous training} approach is used problems may occur influencing the result, if updates are too sporadic and different versions of values drift too far apart.
+To avoid this problem GPU-Cluster configurations do not rely on complete asynchronous training startegies, but on the already described mixed-asynchronous approach, minimizing this problem while still reducing possible communication overhead.
 
-To ease the effect of \emph{GPU to GPU communication},by relieveing the CPU of managing them, technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead.
-Also shared memory between GPUs can come in place saving synchronization between them.
+To ease the effect of \emph{GPU to GPU communication},by relieveing the CPU of managing them, technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU-Cluster or a single system and does not remove all communication overhead.
+Also shared memory between GPUs can be useful, reducing synchronization between them (and enabling, to an extent, model parallelism cross GPUs)
 
@@
 %bandwith between clusters
-\emph{Cluster communication} becomes only more and more concerning if more GPU-Clusters are used, because the connection between them may slower than the interfaces of RAM, GPU and storage used therefore creating a new bottleneck.
-This problem can be reduce if training data is stored by the cluster, which has the downside of further increasing the cost of these setups.
+\emph{Cluster communication} becomes only more and more concerning when more GPU-Clusters are used, because the connection between them may be slower than the interfaces of RAM, GPU and storage used therefore creating a new bottleneck.
+This problem can be reduced if training data is stored by the cluster, which has the downside of further increasing the cost of these setups.
 
@@
 %Summary
-During our survey we observed that the most influencing factors are the basic choice of neural network to be used, this includes its complexity as well as the actual architecture the network is build of.
-Further the processing unit that is chosen has a strong influence on the quality and speed the final network that lies within acceptable error limits will be reached.
+During our survey we observed that the most influencing factors are the basic choice of neural network to be used, this includes its complexity as well as the actual architecture of the network is build of.
+Furthermore the processing unit that is chosen has a strong influence on the quality and speed of training the network.
 Based on the choice of this other factors came into consideration influencing the process, like communication between processing units and parallelism of data between them.
 They are not quite as influential as the first two but can lead to a strong improvement or worsening.
-We observed that even in example systems that handle communication efficiently about 35 \% of the calculation time is sacrificed to it.
+We observed that even in example systems that handle communication efficiently about 35 \% of the calculation time is sacrificed to it, as in our example previously.
 
 %Takeaway
-Especially data transfer time and alike carries a potential to lengthen the training time of an significant amount.
-To ease this technology like GPUDirect\cite{nvidiagpudirect2017} and data prefetching\cite{yang2010gpgpu} can be used.
+Especially data transfer time and alike carries a potential to lengthen the training time to a significant amount.
+To ease this, technologies like GPUDirect\cite{nvidiagpudirect2017} and data prefetching\cite{yang2010gpgpu} can be used.
 It may be possible to decrease the effect of the data flow further by introducing a direct memory transfer unit which can indepently organize synchronization and can handle parallel transfers to and from multiple GPUs.
@@
 Although this could break some security limits because of unauthorized memory accesses.
+
+%End
+In the end, these are factors that need to be considered by both data scientist working with networks, and systems developers who build deep learning systems.
+Considering the benefits of these technologies, we believe that more research on improving these runtimes could be beneficial.
+As future work we suggest that edge computation for networks (a pertinent topic for Intelligent Personal Assistants like Google Now, Siri or Alexa) should also be studied more by system developers.
 

### 2018-01-12 paper.tex
again some changes

@@
 This popularity only increased over the recent years thorugh widely publicly known applications ranging from DeepBlue to IBM Watson and AlphaGo.
-Because of these successes Machine Learning is sometimes viewed as an omnipotent solution to every imaginable problem.
-This is most certainly not the case, based on our studies we can conclude that their capabilities depend on done design and engineering work.
-In this paper we want to give an short introduction into the most performance influencing factors to performance, during the most time intensive phase, the training.
-
+Because of these successes Machine Learning, and the specialized sub-field of Deep Learning,  is sometimes viewed as an omnipotent solution to every imaginable problem.
+This is most certainly not the case.
+Based on our studies we can conclude that the capabilities of neural networks (i.e., the model used in deep learning) depend on the quality of their training.
+This process, in turn, can be time-consuming.
+Its performance depends on good design and engineering work.
+In this paper we want to give an short introduction into the most performance influencing factors when working with neural networks, with a focus on their training stage.
 \end{abstract}
@@
 
-During our study we found no comprehensive study touching this thema not being based on the comparison of frameworks or on an experimental execution on different platforms or frameworks.
+During our research we found no comprehensive study touching this topic.
+The most closely related work are studies with a focus on comparing different platforms or frameworks.
 This paper is based on these publications.
@@
 Afterwards we have a look, on one hand, at small scale neural networks running on a single machine without large data sets and on the other hand, large scale neural networks operating with more layers and nodes, large data sets and a distributed system to handle the computational power required by the neural network training algorithms to be completed in reasonable time.
-We divide our survey into, the choice of neural network, the choice of processing units, including basic data distribution strategies for multi-processing, a discussion of the challenges in running small and large scales.
+We divide our survey into
+\begin{itemize}
+\item the choice of neural network
+\item the choice of processing units, including basic data distribution strategies for multi-processing
+\item a discussion of the challenges in running small and large scales.
+\end{itemize}
 
@@
 \centering
-\includegraphics[width=0.4\textwidth]{neuralnetworks.png}
+\includegraphics[width=0.5\textwidth]{neuralnetworks.png}
 \caption{"A mostly complete chart of Neural Networks" by "The Asimov Institute" extract of only in this paper referenced networks\cite{zoo}}
@@
 
-The term "Deep Learning" itself is just a rebranding of the since the 80s existing concept of multilayered perceptrons.
+The term "Deep Learning" itself is just a rebranding of the since the 50s existing concept of multilayered perceptrons\cite{frank1957perceptron}.
 Its basic idea is to simply use more than 1 layer in a neural network to increase the accuracy of the learned model\cite{deeplearning101}.
 A good defintion of Deep Learning is delivered by Deng in his 2014 book "Deep learning: methods and applications"\cite{deng2014deep}: "A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification".
-Additionally to the reuse of multilayer perceptrons the ideas of Deep Belief Networks and unsupervised pretraining have been introduced.
+Additionally to the reuse of multilayer perceptrons the ideas of Deep Belief Networks and unsupervised pretraining were introduced in 2006\cite{hinton2006fast}, opening the door for continuous improvements in different aspects of neural networks.
+Among the relatively recent developments in neural networks has been the adoption of GPUs for processing, as exemplified in work winning the ImageNet challenge\cite{krizhevsky2012imagenet}.
 
@@
 \centering
-\includegraphics[width=0.4\textwidth]{pop.png}
+\includegraphics[width=0.5\textwidth]{pop.png}
 \caption{Popularity in \% of the term "deep learning" on Google in relation to maximum peak, plotted with the Google Trends tool provided by Google}
@@
 
-Since then Deep Learning has experienced an almost exponential growth in popularity as seen in Figure 2, where the popularity steeply increases around 2014.
+Since then Deep Learning has experienced an almost exponential growth in popularity as seen in Figure 2, where the popularity steeply increased around 2014.
 %%%%END DEEP LEARNING
@@
 The most commonly used shallow Neural Networks are Feed Forward, Auto Encoders and Restricted Boltzmann Machines, which all have only a single layer of hidden nodes and therefore are quite restricted in what they can compute based on the input data.
+They are shown in Figure 1.
 
-Feed Forward Networks(FFN) simply connects input, hidden and output layers with an all to all connection between the layers with the number of outputs being equal or smaller to the number of hidden nodes.
+Feed Forward Networks(FFN) simply connect input, hidden and output layers with an all to all connection between the layers with the number of outputs being equal or smaller to the number of hidden nodes.
 
-Auto Encoders(AE) are of an similar build but the hidden layer compresses the input information into fewer nodes while these are later again decompressed to output nodes that match the previously given input nodes, therefore AE can be classified as a generative network.
+Auto Encoders(AE) are of a similar build but the hidden layer compresses the input information into fewer nodes while these are later again decompressed to output nodes that match the previously given input nodes, therefore AE can be classified as a generative network.
 
@@
 %deep
-A more advanced architecture is to increase the number of hidden layers to enlarge the application possibility of these networks for more complex models these are then called Deep Networks.
-The more commonly used ones are Variational Auto Encoders , Deep Belief Networks, Generative Adversial Networks, Recurrent Neural Networks ,e.g.
+A more advanced architecture is to increase the number of hidden layers to enlarge the application possibility of these networks for more complex models, these are then called Deep Networks.
+The more commonly used ones are Variational Auto Encoders , Deep Belief Networks, Generative Adversial Networks, Recurrent Neural Networks, e.g.
 Long short-term memory networks, Deep Convolutional Neural Networks, Deconvolutional Networks, and Recursive Neural Networks.
+They are also shown in Figure 1.
 
@@
 
-Deep Convolutional Networks(DCN) are based on Convolutional Nodes compressing the incoming information into the needed base model and then calculating the required output.\cite{PattersonGibson17} They are used to abstract input information into more general models.
-These networks especially are being used to process images.
+Deep Convolutional Networks(DCN) are based on Convolutional Nodes compressing the incoming information into the needed base model and then calculating the required output\cite{PattersonGibson17}.
+They are used to abstract input information into more general models.
+These networks are specially being used to process images.
 
@@
 Some Networks like DBN use a slightly
-diverged "Gentle" Backpropagation, which relies on a learning rate after the initial values have been found from the training data with the help of the pretrained RBM \cite{PattersonGibson17}.
+different "Gentle" Backpropagation, which relies on a learning rate after the initial values have been found from the training data with the help of the pretrained RBM \cite{PattersonGibson17}.
 %%%%END TRAINING ALGORITHMS
@@
 
-Further can the required training data exceed direct accessible memory from the processing unit, enhancing the problem of delivering data to all processing units.
-
 Furthermore if the training data exceeds the memory for the processing unit, the problem of delivering data to all units becomes more pressing.
@@
 This decision can have the biggest influence on computation time.
-We divide our processing unit into three groups starting with the CPU,which have a greater memory capability because they can be equipped with additional RAM sticks but possess a smaller parallelization limitation than GPUs because of their different architecture, then the GPU,with a much larger architectural focus on multithreading than CPUs, and following GPU Clusters, being simply a group of GPUs either in a single node or spanning multiple nodes, like they are used in supercomputers.
+We divide our processing unit into three groups starting with the CPUs, which have a greater memory capability because they can be equipped with additional RAM sticks but possess a smaller parallelization limitation than GPUs because of their different architecture, then the GPUs,with a much larger architectural focus on multithreading than CPUs, and following GPU-Clusters, being simply a group of GPUs either in a single node or spanning multiple nodes, like they are used in supercomputers.
 
@@
 %Transition
-Next will follow the choice of neural network and its influence on performance.
+Next will follow the choice of neural netwok as well as the setting of some hyper-parameters like the amount of nodes, followed by the choice of the processing unit
 
@@
 \subsubsection{Shallow Networks}
-Simple task can be done by a network which is not deep, like a FFN or a RBM.
+Simple learning task can be done by a network which is not deep, like a FFN or a RBM.
 They are faster to train and even larger training sets can be completed in a reasonable time without multithreading or usage of GPUs.
@@
 More complex tasks often require a deep neural network that can learn a more complex model than shallow networks and offer a greater accuracy while doing this.
-The major downside of these networks is that they require a larger amount of time to train and can be oversized for the actual learning task and might adding unnecessary computation time.
+The major downside of these networks is that they require a larger amount of time to train and can be oversized for the actual learning task, adding unnecessary computation time.
 Large scale applications mostly contain deep networks because of their larger training sets and more complex models underlying the actual task.
 
-We can not offer a clear answer about when to use which kind of network, every problem must be individually analyzed and based on the information, and complexity, as well as the acceptable error rate of the final network the decision has to be made.
+We can not offer a clear answer about when to use which kind of network.
+Every problem must be individually analyzed and based on the information, and complexity, as well as the acceptable error rate of the final network the decision has to be made.
 For cases that are unclear a shallow network can first be constructed,because of their short training time, and if the result is unsatisfying a more complex network can be considered.
@@
 
-An also important choice is the setting of the initial values of connection weights, here it is advisable to start at low values like $0.1$ to prevent straying too far from the optimal solution, since algorithm like backpropagation mostly find a local optimum initial values too far of from the global optimum will deliver worse results.
+Initial values, next to the number of nodes, the type of network, the activation functions, the activation thresholds, learning rate, the size of the batches, the training algorithm, the number of training epochs, and others, are called the hyper-parameters of the network.
+We can define these, informally, as the configurable factors that determine the training process in the network and the convergence of error rates.
+In our discussion, (given that our focus is on performance, with a data management perspective) we only consider the number of nodes.
+The precise effect of other hyper-parameters is, arguably less data management related, and thus, beyond the scope of our work.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{a.png}
+
+\caption{Time per mini-batch in s compared between different amount of GPUs used\cite{shi2016benchmarking}}
+\label{fig_m_gpu}
+\end{figure}
+
 %speedup figure
@@
 \subsubsection{Synchronization between GPUs}
-For synchronizing Multi-GPU setups like in Figure 4 there are theoretically two choices.
+For synchronizing Multi-GPU setups like in Figure 5 there are theoretically two choices.
 One being the synchronous method in which after a value is calculated it is immediately shared with every other GPU.
@@
 This depends on the network model and training algorithm.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{a.png}
-
-\caption{Time per mini-batch in s compared between different amount of GPUs used\cite{shi2016benchmarking}}
-\label{fig_m_gpu}
-\end{figure}
 
@@
 In doing this we again increase raw computation power, but also increase the total amount of communication required.
-Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in (Figure 3)
+Similar to multiple GPUs this leads at the beginning with 2 or more Clusters to a stronger improvement than later on with 8 to 16 Clusters as seen in (Figure 4)
 
